### 2019-10-29 article.Rnw
Files for first draft of JSS paper

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}

%% another package (only for this demo article)
\usepackage{framed}

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%
<<kintroptions, echo=FALSE>>=
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  cache = TRUE,
  fig.width = 7
)
@
%

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Anita K Nandi\\University of Oxford
   \And Tim CD Lucas\\University of Oxford
   \And Rohan Arambepola\\University of Oxford
   \AND Andre Python\\University of Oxford
   \And Daniel J Weiss\\University of Oxford}
\Plainauthor{Anita K Nandi, Tim CD Lucas, Rohan Arambepola, Andre Python, Daniel J Weiss}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{disaggregation: An \proglang{R} package for Bayesian spatial disaggregation modelling}
\Plaintitle{disaggregation: An R package for Bayesian spatial disaggregation modelling}
\Shorttitle{disaggregation modelling in \proglang{R}}

%% - \Abstract{} almost as usual
\Abstract{
  Disaggregation modelling, or downscaling, has been used across many fields including, epidemiology, ecology and something else. These models use repsonse data agreggated of large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity. This paper presents the \proglang{R} package \emph{disaggregation}, which provides the functionality to streamline the process of running a disaggregation model for fine scale predictions.
}

%% - \Keywords{} with LaTeX markup, at least one required
%% - \Plainkeywords{} without LaTeX markup (if necessary)
%% - Should be comma-separated and in sentence case.
\Keywords{Bayesian spatial modelling, disaggregation modelling, TMB}
\Plainkeywords{Bayesian spatial modelling, disaggregation modelling, TMB}

%% - \Address{} of at least one author
%% - May contain multiple affiliations for each author
%%   (in extra lines, separated by \emph{and}\\).
%% - May contain multiple authors for the same affiliation
%%   (in the same first line, separated by comma).
\Address{
  Anita Nandi\\
  Malaria Atlas Project\\
  Big Data Institute\\
  Old Road Campus\\
  University of Oxford\\
  Oxford, UK\\
  E-mail: \email{anita.nandi@bdi.ox.ac.uk}
}

\begin{document}

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

\section{Introduction} \label{sec:intro}

Methods for estimating high resolution repsonse information from models of aggregated repsonse data over large spatial regions are 
becoming increasingly sought after in disease risk~\cite{li2012log, diggle2013spatial}, expecially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}, and ecological speices distribution~\cite{keil2013downscaling, barwell2014can}. The aggreagtion of repsonse data over large 
heterongenous regions is problematic for understanding the fine scale distribution. However, by using fine scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.

Disaggregation modelling is unorthodox as the predictions are at a different scale to the response data, i.e. the number of rows in the covariate data is different to that of the repsonse data. Traditional spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate}, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem. Disaggregation models can be implemented in \pkg{TMB} (\cite{tmb2016}) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial. The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models with some level of flexibility.

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

\section{Disaggregation modelling} \label{sec:model}

Let's say we have response data for $N$ polygons, which corresponds to count data for the property of interest within that polygon, $y_j$. By count data here we mean the data at the level that can be aggregated by simple addition, i.e. the count data of the polygon is given by the sum of the count data for all the pixels within that polygon. The rate, $p_i$ is defined such that the number of cases in a pixel can be calculated by multiplying the rate by the aggregation raster. For example, in epidemiology, we may have as our response the number of people with a certain disease (case incidence). Our rate would be the incidence rate as a fraction of the population, where the aggregation raster corresponds to population.

For the disaggregation model, we first predict the rate, $pred_i$, at pixel level:
%
\begin{equation} \label{eq:lp}
link(pred_i) = \beta_0 + \beta X + GP(s_i) + u_j
\end{equation}
%

where $X$ are the covariates, $GP$ is the gaussian random field and $u_j$ is the polygon-specific iid random effect. The Gaussian random field has a Matern covariance function, with hyperparameters $\rho = \frac{\sqrt{8\lambda}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating the magntiude of the variation of the field. Each of the model parameters and hyperparameters are given priors, which is discussed later.

The pixel predictions are then aggregated to the polygon level using the weighted sum (via the aggregation raster, $agg_i$):

%
\begin{equation} \label{eq:agreggate1}
cases_j = \sum_{i \epsilon j} pred_i \times agg_i
\end{equation}
\begin{equation} \label{eq:agreggate2}
rate_j = \frac{\sum_{i \epsilon j} pred_i \times agg_i}{\sum_{i \epsilon j} agg_i}
\end{equation}
%

The different likelihoods correspond to slightly different models ($y_j$ is the repsonse count data):

%
\begin{itemize}
\item {\textbf{Gaussian} ($\sigma_j$ is the dispersion of the polygon data)
\begin{equation} \label{eq:gaussian}
dnorm(y_j/\sum agg_i, rate_j, \sigma_j)
\end{equation}
Here $\sigma_j = \sigma \sqrt{\sum agg_i^2} / \sum agg_i$, where  $\sigma$ is the dispersion of the pixel data (a parameter learnt by the model) }

\item \textbf{Binomial} (For a survey in polygon j, $y_j$ is the number positive and $N_j$ is the number tested),
\begin{equation} \label{eq:binomial}
dbinom(y_j, N_j, rate_j)
\end{equation}

\item \textbf{Poisson} (predicts incidence count),
\begin{equation} \label{eq:poisson}
dpois(y_j, cases_j)
\end{equation}

\end{itemize}
%

Note the normal likelihood is modelled in rate space as it is unreasonable to assume that the dispersion of the likelihood is the same for polygons containing different numbers of cases. Also, the binomial likelihood can be used for survey data, for example when the response data consists of the number of people who tested positive for a disease and the total number of people that were tested.

\subsection{Priors}

For each of the model parameters we specify priors and hyperpriors. The regression parameters and intercept are given Gaussian priors, where the default priors are $\beta_0 \sim N(0, 2)$ and $\beta_i \sim N(0, 0.4)$.

The field parameters are given penalised complexity priors, $\rho_{min}$ and $\rho_{prob}$, where $P(\rho < \rho_{min}) = \rho_{prob}$, and the variation, $\sigma_{max}$ and $\sigma_{prob}$, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$. The effect of these hyperpriors is to regularise the field to have smaller variations over larger distances. Similarly for the iid effect, the priors are penalised complexity priors on the standard deviation, $sd_{max}$ and $sd_{prob}$, such that $P(sd > sd_{max}) = sd_{prob}$. This has the effect of regularising the iid effect to be small.

For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(shape = 1, rate = 5e-05)$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values. This is the same as the prior set by INLA for the dispersion of the normal likelihood.

\section{TMB implementation} \label{sec:implementation}

The \pkg{disaggregation} package is built on Template Model Builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for buidling flexible complex models based on C++. \pkg{TMB} combines the packages CppAD~\cite{bell2012cppad}, for automatic differentiation in C++, Eigen~\cite{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\cite{chen2008algorithm}, a package for efficient computation of sparse matrices. The use of these packages allows an efficient implementation of the applied Laplace approximation~\cite{skaug2006automatic} with exact derivatives. \pkg{TMB} calculates first and second order derivatives of the objective function using automatic differentiation~\cite{griewank2008evaluating}.

More about mathematical details of Laplace approximation and automatic differentiation.

The \pkg{disaggregation} package contains a C++ function that defines the model and computes the joint likelihood as a function of the parameters and the random effects, in the format expected by \pkg{TMB}. The \pkg{TMB} package then calculates estimates of both parameters and random effects using the Laplace approximation for the likelihood. Evaluation of the objective function and its derivatives is performed in \proglang{R}.

\section{Package usage} \label{sec:code}

In this section we show how to use the \pkg{disaggregation} package uasing real malaria data of aggregated malaria case counts across Madagascar, and real environmental covariates, such as temperature, rainfall and elevation, to inform spatial hetrogenity in malaria risk. The latest version of \pkg{disaggregation} should always be available from the Comprehensive R Archive Network (CRAN) at \href{http://CRAN.R-project.org/package=disaggregation}{http://CRAN.R-project.org/package=disaggregation}. Run the following commands to install and load the package.

%
<<installrun, echo=FALSE, message=FALSE>>=
#devtools::install_github("aknandi/disaggregation")
library(disaggregation, quietly = TRUE)
library(raster, quietly = TRUE)
library(dplyr, quietly = TRUE)
library(magrittr, quietly = TRUE)
@
%

%
<<installview, eval=FALSE>>=
devtools::install_github("aknandi/disaggregation")
library(disaggregation)
library(raster)
library(dplyr)
library(magrittr)
@
%

We then setup the data to use in the \pkg{disaggregation}.
%
<<setuppaths>>=
API_path <- 'data/api_subnational.csv'
population_path <-'data/pop_at_risk_pf_2013.tif'
shapefile_path <- 'data/admin2019_2.shp'

covariate_raster_paths <- c(
  'data/MODIS_LST_Day.mean.5km.tif',
  'data/MODIS_EVI.mean.5km.tif',
  'data/SRTM_elevation.5km.mean.tif',
  'data/MODIS_LST_Day.SD.5km.tif'
)

# Make a RasterStack of covariates to the extent of Madagascar and scale values
covariates <- lapply(covariate_raster_paths, raster)
covariate_stack <- stack(covariates)
covariate_stack <- crop(covariate_stack, extent(c(42, 52, -26, -10))) %>% scale

# Read in the incidence data and only select data from Madagascar 2013
api <- read.csv(API_path)
api <- api %>% filter(iso3 == 'MDG', year == 2013)

# Read in the shapefile and only keep those that contain incidence data
shapes <- shapefile(shapefile_path)
shapes <- shapes[shapes$ID_2 %in% as.character(api$fdef_id), ]

# Add incidence information to the shapefile (incidence = API * population/1000)
indices <- match(shapes$ID_2, as.character(api$fdef_id))
shapes$inc <- (api$api_mean_pf[indices] * api$pop_at_risk_pf[indices])/1000

# Read in the population raster and crop it to the extent of Madagascar
population_raster <- raster(population_path)
population_raster <- crop(population_raster, extent(c(42, 52, -26, -10)))
@
%

The main functions are \code{prepare_data}, \code{fit_model} and \code{predict}.

Firstly, we use the \code{prepare_data} function to setup all the data in the format needed in the disaggregation modelling. This function performs various data manipulation tasks to create objects that are required to provide for fitting the model. The required input data for the \code{prepare_data} function are a SpatialPolygonDataFrame containing the response data and a RasterStack of covariates to be used in the model. The variable names in the SpatialPolygonDataFrame of the response count data and the polygon ID are defined by the user.

An optional aggregation raster can be provided. The aggregation raster defines how the pixels within each polygon are aggregated. The disaggregation model performs a weighted sum of the pixel prediction, weighted by the pixel values in the aggregation raster. In this case our pixel predictions are malaria incidence rate, so we use the population raster to aggregate pixel incidence rate by summing the number of cases (rate weighted by population). If no aggregation raster is provided a uniform distribution is assumed, i.e. the pixel predictions are aggregated to polygon level by summing the pixel values.

The values of the covariates (as well as the aggregation raster, if given) are extracted at each pixel within the polygons and stored as a data.frame with a row for each pixel and a column for each covariate (\code{parallelExtract} function). The extraction of each covariate is performed in parallel over the number of cores defined by the argument \code{ncores}. The values extracted from the aggregation raster are returned as an array of values, one for each pixel.

In order to know which pixels are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function). Additionally, an INLA mesh is built to use for the spatial field (\code{build_mesh} function). The \code{mesh.args} argument allows you to supply a list of INLA mesh parameters to control the mesh used for the spatial field. The mesh can take a while to construct, so if you want to prepare the data without buidling the mesh, the user can set the \code{makeMesh} flag to FALSE. However, you cannot then fit the disaggregation model without the mesh.

If there are any NAs in the response or covariate data within the polygons the \code{prepare_data} method will return an error. This can be dealt with using the \code{na.action} flag, which is automatically off.  Ideally the NAs in the data would be dealt with by the user beforehand, however, setting na.action = TRUE will automatically deal with NAs. It removes any polygons that have NAs as a response, sets any aggregation pixels with NA to zero and sets covariate NAs pixels to the median value for the that covariate.

%
<<prepare_data>>=
dis_data <- prepare_data(polygon_shapefile = shapes, 
                         covariate_rasters = covariate_stack, 
                         aggregation_raster = population_raster, 
                         mesh.args = list(max.edge = c(0.7, 8), 
                                          cut = 0.05, 
                                          offset = c(1, 2)),
                         id_var ='ID_2',
                         response_var = 'inc',
                         na.action = TRUE,
                         ncores = 8)
@
%

We can see a summary of the data, using the generic \code{summary} function, and \code{plot} the data. The summary function returns information on how many pixels and polygons the data contains, how many pixels in the smallest and largest polygons and a summary of the covariate data. The \code{plot} functions plots a map of the polygon response data, the covariate rasters and the INLA mesh.

%
<<showdata>>=
summary(dis_data)
plot(dis_data)
@
%

The \code{prepare_data} function returns an object of class \code{disag.data}, which is designed to be used directly in the \code{fit_model} function.

Now we can fit the disaggregation model using \code{fit_model}. Here we use a poisson likelihood for the incident count data with a log link function. Options for the likelihood are gaussian, poisson and binomial, and options for the link function are logit, log and identity. The spatial field and iid effect are components of the model by default, they can be turned off using the \code{field} and \code{iid} flags. We specify all of the priors for the model in a single list. Hyperpriors for the field are given as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{min}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{min}) = \sigma_{prob}$. Similarly, you specify penalised complexity priors for the iid effect.
The \code{iterations} argument specifies the maximum number of iterations the model can run for to find an optimal point. In order to print more verbose output you can set the \code{silent} argument to FALSE.

%
<<fit_model>>=
fitted_model <- fit_model(data = dis_data,
                          iterations = 1000,
                          family = 'poisson',
                          link = 'log',
                          priors = list(priormean_intercept = 0,
                                        priorsd_intercept = 2,
                                        priormean_slope = 0.0,
                                        priorsd_slope = 0.4,
                                        prior_rho_min = 3,
                                        prior_rho_prob = 0.01,
                                        prior_sigma_max = 1,
                                        prior_sigma_prob = 0.01))
@
%

We can get a summary and plot of the model output. The \code{summary} function gives the estimate and standard error of the fixed effect parameters in the model, as well as the negative log likelihood and in-sample performance metrics (root mean squared error, mean absolute error, pearson correlation coefficient and spearman rank correlation coefficient). The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions.

%
<<showmodel>>=
summary(fitted_model)
plot(fitted_model)
@
%

The \code{fit_model} function returns an object of class \code{fit.result}, which is designed to be used directly in the \code{predict} function.
Therefore, now once we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.

To predict over a different spatial extent to that used in the model, a RasterStack covering the region to make predictions over can be passed as the \code{newdata} argument. If this argument is not given, predictions are made over the covaraite data used in the fit. If the user wants to include the iid effect from the model as a component in the prediction then the \code{predict_iid} logical flag should be set to TRUE, otherwise, the iid effect will not be predicted.

For the uncertainty calculations, parameter values are sampled from the posterior distribution and summarised. The number of parameter draws used to calculate the uncertainty is set by the user via the \code{N} parameter (default: 100), and the size of the credible interval (75\%, 95\%, ...) to be calculated when summarising is set via the argument \code{CI} (default: 0.95).

%
<<predict_model>>=
model_predictions <- predict(fitted_model)
@
%

The function \code{predict} returns a list of two objects: the mean predictions (of class \code{predictions}) and the uncertainty rasters (of class \code{uncertainty}). The mean predictions contain a raster of the mean prediction of the incidence rate, as well as rasters of the field, iid (if predicted) and covariate component of the linear predictor. The uncertainty contains a RasterStack of the prediction realisations and a RasterStack of the upper and lower credible intervals.

The \code{plot} function can be used on both the \code{predictions} and \code{uncertainty} objects.

%
<<showpredictions>>=
plot(model_predictions$mean_predictions)
plot(model_predictions$uncertainty_predictions)
@

Using three simple functions we have been able to fit a Bayesian spatial disaggregation model and predict pixel-level incidence rate across Madagascar using aggregated incidence data and pixel-level environmental covariates.

\section{Comparison with MCMC} \label{sec:comparison}

In this section we show performance comparisons between modelling using this \pkg{disaggregation} package, based on \pkg{TMB}, and using Markov chain Monte Carlo (MCMC). MCMC algorithms sample the probability space to produce an esimate of the probability distribution. Tese algorithms tend to be slow, as it takes a long time to effectively sample the likelihood space. Using the Laplace approximation, automatic differentiation and efficienct computation of sparse matrices within \pkg{TMB}, the time taken to optimise the model is be considerably shorter. Here we compare the time and performance of the two techniques.

The model described above for malaria in Madagascar is fitted by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm. The model is fitted three different ways: firstly, the MCMC algorithm is run for 1000 iteration, which takes 8 hours; secondly algorithm is then run for 26 iterations, taking 3 minutes; and thirdly the model is optimised using \pkg{TMB} within the \pkg{disaggregation} package, which takes 56 seconds. Fitted parameter values for all three methods are given in the table below.

%
<<mcmc, eval=FALSE>>=
library(tmbstan)

m <- fit_model(data = dis_data, 
               iterations = 1,
               family = 'poisson', 
               link = 'log',
               priors = list(priormean_intercept = 0,
                             priorsd_intercept = 2,
                             priormean_slope = 0.0,
                             priorsd_slope = 0.4,
                             prior_rho_min = 3,
                             prior_rho_prob = 0.01,
                             prior_sigma_max = 1,
                             prior_sigma_prob = 0.01))

# Running the MCMC algorithm for 3 minutes
start <- Sys.time()
mcmc_out <- tmbstan(m$obj, chains = 4, iter = 30, 
                    cores = getOption('mc.cores', 4))
end <- Sys.time()
print(end - start)

# Running the MCMC algorithm for 8 hours
start <- Sys.time()
mcmc_out2 <- tmbstan(m$obj, chains = 4, iter = 1000, 
                     cores = getOption('mc.cores', 4))
end <- Sys.time()
print(end - start)

# Plot the trace of the parameter values for each sampling method
stan_trace(mcmc_out, pars = c('intercept', 
                              'slope[1]', 'slope[2]', 'slope[3]', 'slope[4]', 
                              'iideffect_log_tau', 'log_sigma', 'log_rho'))
stan_trace(mcmc_out2, pars = c('intercept', 
                               'slope[1]', 'slope[2]', 'slope[3]', 'slope[4]', 
                               'iideffect_log_tau', 'log_sigma', 'log_rho'))
@
%

The trace of the MCMC parameter values for the long and the short experiments is given in Figures~\ref{fig:mcmcshort} and \ref{fig:mcmclong} respectively. It can be seen that running the MCMC algorithm for 3 minutes does not show mixing between the different chain; the algorithm much be run for much longer, for exmaple 8 hours, to get sufficient chain mixing. The \pkg{disaggregation} package produces similar results to the MCMC algorithm that runs for 8 hours, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds. Therefore, it is clear that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.

%
\begin{table}
\begin{tabular}{l|ll|ll|ll}
 & \multicolumn{2}{|c|}{MCMC (8 hours)} & \multicolumn{2}{|c|}{MCMC (24 seconds)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
Parameter & Mean & sd & Mean & sd & Mean & sd \\
Intercept & -3.16 & 0.19 & -0.20 & 1.33 & -3.12 & 0.27 \\
Slope 1 & 0.32 & 0.23 & -0.27 & 0.80 & 0.24 & 0.27 \\
Slope 2 & 0.11 & 0.22 & -0.65 & 0.86 & 0.30 & 0.21\\
Slope 3 & -0.55 & 0.18 & -0.35 & 0.95 & -0.37 & 0.20 \\
Slope 4 & -0.23 & 0.16 & 0.12 & 0.35 & -0.24 & 0.21 \\
iid $log(\tau)$ & 1.06 & 0.26 & 0.04 & 1.16 & 1.10 & 0.27 \\
$log(\sigma)$ & -3.29 & 0.61 & -0.87 & 0.94 & -3.55 & 0.59 \\
$log(\rho)$ & 0.75 & 0.34 & 0.65 & 0.93 & -0.56 & 0.30 \\
\end{tabular}
\end{table}
%

%
\begin{figure}[t!]
\centering
\includegraphics[width=\linewidth]{figures/mcmc_short_trace.pdf}
\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 3 minutes.}
\label{fig:mcmcshort}
\end{figure}
%

%
\begin{figure}[t!]
\centering
\includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 8 hours.}
\label{fig:mcmclong}
\end{figure}
%

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Conclusions} \label{sec:conclusions}

Disaggregation modelling, which involves predicting at fine-scale resolution using areal data over heterogenous regions, has become widely used in fields such as epidemiology, ecology and something. The \pkg{disaggregation} package implements Bayesian spatial disaggregation modelling with a simple, easy to use \proglang{R} interface. The package includes simple data preparation, fitting and prediction functions that allow some user defined model flexibility. In this paper we have presented an application of the package, predicting malaria incidence rate across Madagascar from aggregated count data and environmental covariates.

The modelling framework is implemented using the Laplace approximation and automatic differentiation within the \pkg{TMB} package. This allows fast, optimised calculations in C++. These disaggregation models are computationally intensive and take a long time using Markov Chian Monte Carlo (MCMC) optimisation techniques. Using \pkg{TMB}, the models are much faster and produce similar results

The \pkg{disaggregation} package can be extended to include spatio-temporal disaggregation models. This would require a spatio-temporal field as well as dynamic covariates. However, this would be significantly more computationally intensive. Additionally, tools for cross-validation could be included within the package. Cross-validation of spatial models is non-trivial as random cross validation over estimates the performance of the model due to spatial autocorrelation and spatial cross-validation will show good performance if your underlying model is generalisable to new geographic regions.

%% -- Optional special unnumbered sections -------------------------------------

\section*{Computational details}

\begin{leftbar}
If necessary or useful, information about certain computational details
such as version numbers, operating systems, or compilers could be included
in an unnumbered section. Also, auxiliary packages (say, for visualizations,
maps, tables, \dots) that are not cited in the main text can be credited here.
\end{leftbar}

The results in this paper were obtained using
\proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
and all packages used are available from the Comprehensive
\proglang{R} Archive Network (CRAN) at
\url{https://CRAN.R-project.org/}.

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2019-10-30 article.Rnw
Tidy up and update MCMC results

@@
 
-More about mathematical details of Laplace approximation and automatic differentiation.
-
 The \pkg{disaggregation} package contains a C++ function that defines the model and computes the joint likelihood as a function of the parameters and the random effects, in the format expected by \pkg{TMB}.
@@
 library(disaggregation, quietly = TRUE)
-library(raster, quietly = TRUE)
-library(dplyr, quietly = TRUE)
-library(magrittr, quietly = TRUE)
 @
@@
 library(disaggregation)
-library(raster)
-library(dplyr)
-library(magrittr)
 @
@@
 
-We then setup the data to use in the \pkg{disaggregation}.
+We then setup the data to use in the \pkg{disaggregation} package.
+
 %
 <<setuppaths>>=
+library(raster)
+
 API_path <- 'data/api_subnational.csv'
@@
   'data/SRTM_elevation.5km.mean.tif',
-  'data/MODIS_LST_Day.SD.5km.tif'
-)
+  'data/MODIS_LST_Day.SD.5km.tif')
 
@@
 covariate_stack <- stack(covariates)
-covariate_stack <- crop(covariate_stack, extent(c(42, 52, -26, -10))) %>% scale
+covariate_stack <- scale(crop(covariate_stack, extent(c(42, 52, -26, -10))))
 
@@
 api <- read.csv(API_path)
-api <- api %>% filter(iso3 == 'MDG', year == 2013)
+api <- api[api$iso3 == 'MDG' & api$year == 2013, ]
 
@@
 \begin{tabular}{l|ll|ll|ll}
- & \multicolumn{2}{|c|}{MCMC (8 hours)} & \multicolumn{2}{|c|}{MCMC (24 seconds)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
+ & \multicolumn{2}{|c|}{MCMC (8 hours)} & \multicolumn{2}{|c|}{MCMC (3 minutes)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
 Parameter & Mean & sd & Mean & sd & Mean & sd \\
-Intercept & -3.16 & 0.19 & -0.20 & 1.33 & -3.12 & 0.27 \\
-Slope 1 & 0.32 & 0.23 & -0.27 & 0.80 & 0.24 & 0.27 \\
-Slope 2 & 0.11 & 0.22 & -0.65 & 0.86 & 0.30 & 0.21\\
-Slope 3 & -0.55 & 0.18 & -0.35 & 0.95 & -0.37 & 0.20 \\
-Slope 4 & -0.23 & 0.16 & 0.12 & 0.35 & -0.24 & 0.21 \\
-iid $log(\tau)$ & 1.06 & 0.26 & 0.04 & 1.16 & 1.10 & 0.27 \\
-$log(\sigma)$ & -3.29 & 0.61 & -0.87 & 0.94 & -3.55 & 0.59 \\
-$log(\rho)$ & 0.75 & 0.34 & 0.65 & 0.93 & -0.56 & 0.30 \\
+Intercept & -3.07 & 0.37 & -3.31 & 0.19 & -3.12 & 0.27 \\
+Slope 1 & 0.22 & 0.27 & 0.50 & 0.43 & 0.24 & 0.27 \\
+Slope 2 & 0.30 & 0.20 & 0.21 & 0.42 & 0.30 & 0.21\\
+Slope 3 & -0.35 & 0.21 & -0.50 & 0.33 & -0.37 & 0.20 \\
+Slope 4 & -0.22 & 0.21 & -0.24 & 0.15 & -0.24 & 0.21 \\
+iid $log(\tau)$ & 1.06 & 0.26 & -0.14 & 1.13 & 1.10 & 0.27 \\
+$log(\sigma)$ & -3.29 & 0.61 & -3.37 & 2.35 & -3.55 & 0.59 \\
+$log(\rho)$ & 0.75 & 0.34 & 2.78 & 1.72 & -0.56 & 0.30 \\
 \end{tabular}

### 2019-11-14 article.Rnw
Update MCMC section with longer run (still not enough). A bit about malaria. Other minor changes

@@
    \And Tim CD Lucas\\University of Oxford
-   \And Rohan Arambepola\\University of Oxford
-   \AND Andre Python\\University of Oxford
-   \And Daniel J Weiss\\University of Oxford}
+   \And Rohan Arambepola\\University of Oxford}
+   %\AND Andre Python\\University of Oxford
+   %\And Daniel J Weiss\\University of Oxford}
 \Plainauthor{Anita K Nandi, Tim CD Lucas, Rohan Arambepola, Andre Python, Daniel J Weiss}
@@
 \Abstract{
-  Disaggregation modelling, or downscaling, has been used across many fields including, epidemiology, ecology and something else.
+  Disaggregation modelling, or downscaling, has been used across many fields including epidemiology and ecology.
 These models use repsonse data agreggated of large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
@@
 However, by using fine scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
+This method has recently been applied in the latest global high-resolution maps of \emph{Plasmodium falciparum} and \emph{Plasmodium vivax} malaria~\cite{falciparum2019, vivax2019}.
 
@@
 Traditional spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate}, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem.
-Disaggregation models can be implemented in \pkg{TMB} (\cite{tmb2016}) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
-The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models with some level of flexibility.
+Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
+The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models with some degree of flexibility.
 
@@
 
-In this section we show how to use the \pkg{disaggregation} package uasing real malaria data of aggregated malaria case counts across Madagascar, and real environmental covariates, such as temperature, rainfall and elevation, to inform spatial hetrogenity in malaria risk.
+In this section we show how to use the \pkg{disaggregation} package using real malaria data of aggregated malaria case counts across Madagascar.
+Malaria is a infectious disease caused by parasites of the \emph{Plasmodium} group, transmitted by \emph{Anopheles} mosquitoes.
+Malaria transmission is therefore closely related to mosquito and parasite development.
+Environmental factors such as temperature, rainfall and elevation have been shown to have signifcant effect on mosquito survival and development; in general mosquitoes favour warm, humid environments with moderate rainfall.
+Therefore, such environmental covariates would be useful in a malaria disaggregation model to inform fine-scale distributions.
+In this model we use the environmental covariates of mean land surface temperature (LST), variation in LST, elevation, and enhanced vegetation index (EVI) to inform spatial hetrogenity in malaria risk.
+These covaraites are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
+
 The latest version of \pkg{disaggregation} should always be available from the Comprehensive R Archive Network (CRAN) at \href{http://CRAN.R-project.org/package=disaggregation}{http://CRAN.R-project.org/package=disaggregation}.
@@
 MCMC algorithms sample the probability space to produce an esimate of the probability distribution.
-Tese algorithms tend to be slow, as it takes a long time to effectively sample the likelihood space.
-Using the Laplace approximation, automatic differentiation and efficienct computation of sparse matrices within \pkg{TMB}, the time taken to optimise the model is be considerably shorter.
+These algorithms tend to be slow, as it takes a long time to effectively sample the likelihood space.
+In contrast, the \pkg{disaggregation} package, which uses the Laplace approximation, automatic differentiation and efficient computation of sparse matrices within \pkg{TMB}, takes a considerably shorter time to optimise the model.
 Here we compare the time and performance of the two techniques.
 
-The model described above for malaria in Madagascar is fitted by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm.
-The model is fitted three different ways: firstly, the MCMC algorithm is run for 1000 iteration, which takes 8 hours; secondly algorithm is then run for 26 iterations, taking 3 minutes; and thirdly the model is optimised using \pkg{TMB} within the \pkg{disaggregation} package, which takes 56 seconds.
-Fitted parameter values for all three methods are given in the table below.
+The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package.
+Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm.
+It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
+The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took x hours.
+This number of iterations was chosen to be the minimum required for the value of Rhat to drop below 1.05 for all model parameters.
+The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
+Fitted parameter values for both of these methods are given in the table below.
 
@@
 
-# Running the MCMC algorithm for 3 minutes
-start <- Sys.time()
-mcmc_out <- tmbstan(m$obj, chains = 4, iter = 30, 
-                    cores = getOption('mc.cores', 4))
-end <- Sys.time()
-print(end - start)
-
 # Running the MCMC algorithm for 8 hours
 start <- Sys.time()
-mcmc_out2 <- tmbstan(m$obj, chains = 4, iter = 1000, 
+mcmc_out <- tmbstan(m$obj, chains = 4, iter = 8000, warmup = 2000,
                      cores = getOption('mc.cores', 4))
@@
                                'iideffect_log_tau', 'log_sigma', 'log_rho'))
-stan_trace(mcmc_out2, pars = c('intercept', 
-                               'slope[1]', 'slope[2]', 'slope[3]', 'slope[4]', 
-                               'iideffect_log_tau', 'log_sigma', 'log_rho'))
 @
@@
 
-The trace of the MCMC parameter values for the long and the short experiments is given in Figures~\ref{fig:mcmcshort} and \ref{fig:mcmclong} respectively.
-It can be seen that running the MCMC algorithm for 3 minutes does not show mixing between the different chain; the algorithm much be run for much longer, for exmaple 8 hours, to get sufficient chain mixing.
-The \pkg{disaggregation} package produces similar results to the MCMC algorithm that runs for 8 hours, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
+The trace of the MCMC parameter values is given in Figures~\ref{fig:mcmclong}.
+It can be seen that the MCMC algorithm has been run for long enough to get sufficient chain mixing.
+The \pkg{disaggregation} package produces similar results to the MCMC algorithm, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
 Therefore, it is clear that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
@@
 \begin{table}
-\begin{tabular}{l|ll|ll|ll}
- & \multicolumn{2}{|c|}{MCMC (8 hours)} & \multicolumn{2}{|c|}{MCMC (3 minutes)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
-Parameter & Mean & sd & Mean & sd & Mean & sd \\
-Intercept & -3.07 & 0.37 & -3.31 & 0.19 & -3.12 & 0.27 \\
-Slope 1 & 0.22 & 0.27 & 0.50 & 0.43 & 0.24 & 0.27 \\
-Slope 2 & 0.30 & 0.20 & 0.21 & 0.42 & 0.30 & 0.21\\
-Slope 3 & -0.35 & 0.21 & -0.50 & 0.33 & -0.37 & 0.20 \\
-Slope 4 & -0.22 & 0.21 & -0.24 & 0.15 & -0.24 & 0.21 \\
-iid $log(\tau)$ & 1.06 & 0.26 & -0.14 & 1.13 & 1.10 & 0.27 \\
-$log(\sigma)$ & -3.29 & 0.61 & -3.37 & 2.35 & -3.55 & 0.59 \\
-$log(\rho)$ & 0.75 & 0.34 & 2.78 & 1.72 & -0.56 & 0.30 \\
+\centering
+\begin{tabular}{l|ll|ll}
+ & \multicolumn{2}{|c|}{MCMC (x hours - TBC)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
+Parameter & Mean & SD & Mean & SD \\
+\hline
+Intercept & -3.06 & 0.42 & -3.12 & 0.27 \\
+Slope 1 & 0.17 & 0.29 & 0.24 & 0.27 \\
+Slope 2 & 0.31 & 0.22 & 0.30 & 0.21\\
+Slope 3 & -0.34 & 0.22 & -0.37 & 0.20 \\
+Slope 4 & -0.24 & 0.26 & -0.24 & 0.21 \\
+iid $log(\tau)$ & 1.11 & 0.28 & 1.10 & 0.27 \\
+$log(\sigma)$ & -3.35 & 0.60 & -3.55 & 0.59 \\
+$log(\rho)$ & 0.74 & 0.34 & -0.56 & 0.30 \\
 \end{tabular}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/mcmc_short_trace.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 3 minutes.}
-\label{fig:mcmcshort}
-\end{figure}
-%
-
-%
-\begin{figure}[t!]
-\centering
-\includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 8 hours.}
+\includegraphics[width=\linewidth]{figures/mcmc_long.pdf}
+\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for x hours.}
 \label{fig:mcmclong}
@@
 
-\begin{leftbar}
-If necessary or useful, information about certain computational details
-such as version numbers, operating systems, or compilers could be included
-in an unnumbered section. Also, auxiliary packages (say, for visualizations,
-maps, tables, \dots) that are not cited in the main text can be credited here.
-\end{leftbar}
-
 The results in this paper were obtained using
@@
 \proglang{R} Archive Network (CRAN) at
-\url{https://CRAN.R-project.org/}.
+\url{https://CRAN.R-project.org/}, apart from \pkg{INLA}, which can be 
+installed in \proglang{R} using the command:
+%
+<<inla, eval=FALSE>>=
+install.packages("INLA", 
+                 repos = c(getOption("repos"), 
+                           INLA="https://inla.r-inla-download.org/R/stable"), 
+                 dep=TRUE)
+@
+%
 
@@
 
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
+We would like to thank the Bill and Melinda Gates Foundation for funding this research.
 
@@
 
-\newpage
-
-\begin{appendix}
-
-\section{More technical details} \label{app:technical}
-
-\begin{leftbar}
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
-
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-  \item Title vs.\ sentence case.
-  \item Graphics formatting.
-  \item Naming conventions.
-  \item Turning JSS manuscripts into \proglang{R} package vignettes.
-  \item Trouble shooting.
-  \item Many other potentially helpful details\dots
-\end{itemize}
-\end{leftbar}
-
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-    be used in the references.
-  \item Titles should be in title case.
-  \item Journal titles should not be abbreviated and in title case.
-  \item DOIs should be included where available.
-  \item Software should be properly cited as well. For \proglang{R} packages
-    \code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
-
-\end{appendix}
+% \newpage
+% 
+% \begin{appendix}
+% 
+% \section{More technical details} \label{app:technical}
+% 
+% \begin{leftbar}
+% Appendices can be included after the bibliography (with a page break). Each
+% section within the appendix should have a proper section title (rather than
+% just \emph{Appendix}).
+% 
+% For more technical style details, please check out JSS's style FAQ at
+% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
+% which includes the following topics:
+% \begin{itemize}
+%   \item Title vs.\ sentence case.
+%   \item Graphics formatting.
+%   \item Naming conventions.
+%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
+%   \item Trouble shooting.
+%   \item Many other potentially helpful details\dots
+% \end{itemize}
+% \end{leftbar}
+% 
+% 
+% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
+% 
+% \begin{leftbar}
+% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
+% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
+% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
+% formats of author-year citations and allow to include additional details (e.g.,
+% pages, chapters, \dots) in brackets. In case you are not familiar with these
+% commands see the JSS style FAQ for details.
+% 
+% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
+% when acquiring the entries automatically from mixed online sources. However,
+% it is important that informations are complete and presented in a consistent
+% style to avoid confusions. JSS requires the following format.
+% \begin{itemize}
+%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
+%     be used in the references.
+%   \item Titles should be in title case.
+%   \item Journal titles should not be abbreviated and in title case.
+%   \item DOIs should be included where available.
+%   \item Software should be properly cited as well. For \proglang{R} packages
+%     \code{citation("pkgname")} typically provides a good starting point.
+% \end{itemize}
+% \end{leftbar}
+% 
+% \end{appendix}
 

### 2019-11-18 article.Rnw
Add latest MCMC run to paper. Delete unnecessary plots

@@
 The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package.
-Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm.
+Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm using four chains.
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
-The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took x hours.
-This number of iterations was chosen to be the minimum required for the value of Rhat to drop below 1.05 for all model parameters.
+The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 94 hours.
+This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations and doubling the number of iterations each time, until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
 The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
@@
 
-# Running the MCMC algorithm for 8 hours
+# Running the MCMC algorithm for 94 hours
 start <- Sys.time()
@@
 \begin{tabular}{l|ll|ll}
- & \multicolumn{2}{|c|}{MCMC (x hours - TBC)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
+ & \multicolumn{2}{|c|}{MCMC (94 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
 Parameter & Mean & SD & Mean & SD \\
 \hline
-Intercept & -3.06 & 0.42 & -3.12 & 0.27 \\
-Slope 1 & 0.17 & 0.29 & 0.24 & 0.27 \\
-Slope 2 & 0.31 & 0.22 & 0.30 & 0.21\\
-Slope 3 & -0.34 & 0.22 & -0.37 & 0.20 \\
-Slope 4 & -0.24 & 0.26 & -0.24 & 0.21 \\
-iid $log(\tau)$ & 1.11 & 0.28 & 1.10 & 0.27 \\
-$log(\sigma)$ & -3.35 & 0.60 & -3.55 & 0.59 \\
-$log(\rho)$ & 0.74 & 0.34 & -0.56 & 0.30 \\
+Intercept & -3.09 & 0.33 & -3.12 & 0.27 \\
+Slope 1 & 0.20 & 0.28 & 0.24 & 0.27 \\
+Slope 2 & 0.32 & 0.21 & 0.30 & 0.21\\
+Slope 3 & -0.36 & 0.20 & -0.37 & 0.20 \\
+Slope 4 & -0.24 & 0.23 & -0.24 & 0.21 \\
+iid $log(\tau)$ & 1.08 & 0.27 & 1.10 & 0.27 \\
+$log(\sigma)$ & -3.34 & 0.59 & -3.55 & 0.59 \\
+$log(\rho)$ & 0.72 & 0.31 & -0.56 & 0.30 \\
 \end{tabular}
@@
 \centering
-\includegraphics[width=\linewidth]{figures/mcmc_long.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for x hours.}
+\includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
+\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 94 hours.}
 \label{fig:mcmclong}

### 2019-11-18 article.Rnw
Add caption to table
Minor update to abstract

@@
   Disaggregation modelling, or downscaling, has been used across many fields including epidemiology and ecology.
-These models use repsonse data agreggated of large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
+In these fields surveillance data, aggregated over large regions, is becoming more common, leading to an increasing demand for modelling frameworks that can deal with this data to understand spatial patterns.
+Disaggregation regression models use repsonse data agreggated over large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
 This paper presents the \proglang{R} package \emph{disaggregation}, which provides the functionality to streamline the process of running a disaggregation model for fine scale predictions.
@@
 Methods for estimating high resolution repsonse information from models of aggregated repsonse data over large spatial regions are 
-becoming increasingly sought after in disease risk~\cite{li2012log, diggle2013spatial}, expecially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}, and ecological speices distribution~\cite{keil2013downscaling, barwell2014can}.
+becoming increasingly sought after in disease risk~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}, and ecological speices distribution~\cite{keil2013downscaling, barwell2014can}.
 The aggreagtion of repsonse data over large
@@
 
-Let's say we have response data for $N$ polygons, which corresponds to count data for the property of interest within that polygon, $y_j$.
+Suppose we have response data for $N$ polygons, which corresponds to count data for the property of interest within that polygon, $y_j$.
 By count data here we mean the data at the level that can be aggregated by simple addition, i.e.
 the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
-The rate, $p_i$ is defined such that the number of cases in a pixel can be calculated by multiplying the rate by the aggregation raster.
+The rate, $p_i$, is defined such that the number of cases in a pixel can be calculated by multiplying the rate by the aggregation raster.
 For example, in epidemiology, we may have as our response the number of people with a certain disease (case incidence).
@@
 The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
-Fitted parameter values for both of these methods are given in the table below.
+Fitted parameter values for both of these methods are given in Table~\ref{tab:mcmcresults}.
 
@@
 
-The trace of the MCMC parameter values is given in Figures~\ref{fig:mcmclong}.
-It can be seen that the MCMC algorithm has been run for long enough to get sufficient chain mixing.
-The \pkg{disaggregation} package produces similar results to the MCMC algorithm, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
-Therefore, it is clear that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
-
 %
@@
 \end{tabular}
+\caption{Fitted model parameter values using both MCMC and using \pkg{TMB} within the \pkg{disaggregation} package.}
+\label{tab:mcmcresults}
 \end{table}
 %
+
+The trace of the MCMC parameter values is given in Figures~\ref{fig:mcmclong}.
+It can be seen that the MCMC algorithm has been run for long enough to get sufficient chain mixing.
+The \pkg{disaggregation} package produces similar results to the MCMC algorithm, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
+Therefore, it is clear that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
 

### 2019-11-18 article.Rnw
Minor updates for first draft

@@
 
-where $X$ are the covariates, $GP$ is the gaussian random field and $u_j$ is the polygon-specific iid random effect.
+where $\beta$ are the regression coefficients, $X$ are the covariates, $GP$ is the gaussian random field and $u_j$ is the polygon-specific iid random effect.
+The user defined link function is given by $link$.
 The Gaussian random field has a Matern covariance function, with hyperparameters $\rho = \frac{\sqrt{8\lambda}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating the magntiude of the variation of the field.
@@
 \begin{equation} \label{eq:gaussian}
-dnorm(y_j/\sum agg_i, rate_j, \sigma_j)
+N(y_j/\sum agg_i, rate_j, \sigma_j)
 \end{equation}
@@
 \begin{equation} \label{eq:binomial}
-dbinom(y_j, N_j, rate_j)
+Binom(y_j, N_j, rate_j)
 \end{equation}
@@
 \begin{equation} \label{eq:poisson}
-dpois(y_j, cases_j)
+Pois(y_j, cases_j)
 \end{equation}
@@
 The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 94 hours.
-This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations and doubling the number of iterations each time, until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
+This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations, doubling the number of iterations each time until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
 The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
@@
 
-The trace of the MCMC parameter values is given in Figures~\ref{fig:mcmclong}.
+The trace of the MCMC parameter values is given in Figure~\ref{fig:mcmclong}.
 It can be seen that the MCMC algorithm has been run for long enough to get sufficient chain mixing.
 The \pkg{disaggregation} package produces similar results to the MCMC algorithm, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
-Therefore, it is clear that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
+Therefore, it can be seen that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
 
@@
 
-Disaggregation modelling, which involves predicting at fine-scale resolution using areal data over heterogenous regions, has become widely used in fields such as epidemiology, ecology and something.
+Disaggregation modelling, which involves predicting at fine-scale resolution using areal data over heterogenous regions, has become widely used in fields such as epidemiology and ecology.
 The \pkg{disaggregation} package implements Bayesian spatial disaggregation modelling with a simple, easy to use \proglang{R} interface.
@@
 Additionally, tools for cross-validation could be included within the package.
-Cross-validation of spatial models is non-trivial as random cross validation over estimates the performance of the model due to spatial autocorrelation and spatial cross-validation will show good performance if your underlying model is generalisable to new geographic regions.
+Cross-validation of spatial models is non-trivial due to the spatial autocorrelation in the data.
+
+The \pkg{disaggregation} package provides a simple, useful interface to perform spatial disaggregation modelling, with reasonable flexibility, as well as having the scope to be extended to more complex disaggregation models.
 
@@
 \proglang{R}~\Sexpr{paste(R.Version()[6:7], collapse = ".")} with the
-\pkg{MASS}~\Sexpr{packageVersion("MASS")} package. \proglang{R} itself
+\pkg{TMB}~\Sexpr{packageVersion("TMB")} package. \proglang{R} itself
 and all packages used are available from the Comprehensive

### 2019-11-21 article.Rnw
I would focus it on epi. Its not common in eco and havent heard of it elsewhere.

@@
 \Abstract{
-  Disaggregation modelling, or downscaling, has been used across many fields including epidemiology and ecology.
-In these fields surveillance data, aggregated over large regions, is becoming more common, leading to an increasing demand for modelling frameworks that can deal with this data to understand spatial patterns.
-Disaggregation regression models use repsonse data agreggated over large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
+  Disaggregation modelling, or downscaling, has become an important discipline in epidemiology.
+Surveillance data, aggregated over large regions, is becoming more common, leading to an increasing demand for modelling frameworks that can deal with this data to understand spatial patterns.
+Disaggregation regression models use response data agreggated over large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
 This paper presents the \proglang{R} package \emph{disaggregation}, which provides the functionality to streamline the process of running a disaggregation model for fine scale predictions.

### 2019-11-21 article.Rnw
Just seems better wording maybe.

@@
 Surveillance data, aggregated over large regions, is becoming more common, leading to an increasing demand for modelling frameworks that can deal with this data to understand spatial patterns.
-Disaggregation regression models use response data agreggated over large heterogenous regions and predict the response at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
-This paper presents the \proglang{R} package \emph{disaggregation}, which provides the functionality to streamline the process of running a disaggregation model for fine scale predictions.
+Disaggregation regression models use response data aggregated over large heterogenous regions to make predictions at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
+This paper presents the \proglang{R} package \emph{disaggregation}, which provides functionality to streamline the process of running a disaggregation model for fine-scale predictions.
 }

### 2019-11-21 article.Rnw
I think focus on epi, but say the methods came from eco. Seems a good balance.

@@
 
-Methods for estimating high resolution repsonse information from models of aggregated repsonse data over large spatial regions are 
-becoming increasingly sought after in disease risk~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}, and ecological speices distribution~\cite{keil2013downscaling, barwell2014can}.
-The aggreagtion of repsonse data over large
-heterongenous regions is problematic for understanding the fine scale distribution.
-However, by using fine scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
+Methods for estimating high-resolution risk maps from aggregated response data over large spatial regions are 
+becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}. 
+and ecological species distribution~\cite{keil2013downscaling, barwell2014can}. The aggregation of response data over large 
+heterogenous regions is problematic for understanding the fine-scale distribution.
+However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
 This method has recently been applied in the latest global high-resolution maps of \emph{Plasmodium falciparum} and \emph{Plasmodium vivax} malaria~\cite{falciparum2019, vivax2019}.

### 2019-11-21 article.Rnw
That alegana paper doesnt actually do downscaling. It fits at polygon and just predicts down. Wilson paper is good.

@@
 Methods for estimating high-resolution risk maps from aggregated response data over large spatial regions are 
-becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, alegana2016advances}. 
+becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, wilson2017pointless}. 
 and ecological species distribution~\cite{keil2013downscaling, barwell2014can}. The aggregation of response data over large 

### 2019-11-21 article.Rnw
Oops, wilson paper isnt malaiar. This other sturrock one is a call to arms.

@@
 Methods for estimating high-resolution risk maps from aggregated response data over large spatial regions are 
-becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial}, especially in malaria mapping~\cite{sturrock2014fine, wilson2017pointless}. 
+becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\cite{sturrock2014fine, sturrock2016mapping}. 
 and ecological species distribution~\cite{keil2013downscaling, barwell2014can}. The aggregation of response data over large 

### 2019-11-21 article.Rnw
Not quite sure. The full sentence on our papers seemed ott. Saying it was developed in eco but is now important, not just in methods papers, in epi, works well... maybe.

@@
 becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\cite{sturrock2014fine, sturrock2016mapping}. 
-and ecological species distribution~\cite{keil2013downscaling, barwell2014can}. The aggregation of response data over large 
-heterogenous regions is problematic for understanding the fine-scale distribution.
+Disaggregation regression, was first applied to ecological species distribution modelling~\cite{keil2013downscaling, barwell2014can} but has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
+The aggregation of response data over large heterogenous regions is problematic for understanding the fine-scale distribution.
 However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
-This method has recently been applied in the latest global high-resolution maps of \emph{Plasmodium falciparum} and \emph{Plasmodium vivax} malaria~\cite{falciparum2019, vivax2019}.
 

### 2019-11-21 article.Rnw
Try that sentence again.

@@
 becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\cite{sturrock2014fine, sturrock2016mapping}. 
-Disaggregation regression, was first applied to ecological species distribution modelling~\cite{keil2013downscaling, barwell2014can} but has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
+Disaggregation regression, first applied in species distribution modelling in ecology~\cite{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
 The aggregation of response data over large heterogenous regions is problematic for understanding the fine-scale distribution.

### 2019-11-21 article.Rnw
Tweak.
Were getting stick for this in the point-polygon paper so best to hammer the exact point home wherever we can.

@@
 Disaggregation regression, first applied in species distribution modelling in ecology~\cite{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
-The aggregation of response data over large heterogenous regions is problematic for understanding the fine-scale distribution.
+The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions.
 However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
@@
 Disaggregation modelling is unorthodox as the predictions are at a different scale to the response data, i.e.
-the number of rows in the covariate data is different to that of the repsonse data.
-Traditional spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate}, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem.
+the number of rows in the covariate data is different to that of the response data.
+The spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate}, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \cite{wilson2017pointless}.
 Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.

### 2019-11-21 article.Rnw
Not sure about this one...

@@
 Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
-The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models with some degree of flexibility.
+The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models at the expense of some flexibility.
 

### 2019-11-21 article.Rnw
tweak.
Hopefully clearer.

@@
 
-Suppose we have response data for $N$ polygons, which corresponds to count data for the property of interest within that polygon, $y_j$.
-By count data here we mean the data at the level that can be aggregated by simple addition, i.e.
+Suppose we have response data, $y_j$, for $N$ polygons, which corresponds to count data for the property of interest within that polygon. 
+The process that is being counted occurs in continuous space that we model as a high-resolution, square lattice for convenience.
+The data, $y_j$,  are assumed to be created by the aggregation of the counts over the polygon, i.e.
 the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
 The rate, $p_i$, is defined such that the number of cases in a pixel can be calculated by multiplying the rate by the aggregation raster.
-For example, in epidemiology, we may have as our response the number of people with a certain disease (case incidence).
-Our rate would be the incidence rate as a fraction of the population, where the aggregation raster corresponds to population.
+For example, in epidemiology, we may have as our response the number of people that contract a certain disease in a given period of time (case incidence).
+Our rate would be the incidence rate of cases per population, where the aggregation raster corresponds to population.
 

### 2019-11-21 article.Rnw
This is kinda subtle. You often talk about predicting at the pixel level before we are even talking about estimation. I think its better to just write a model for the process, then estimate theparameters after.
Just weird to start talking about priors without stating that were doing bayes.

@@
 
-For the disaggregation model, we first predict the rate, $pred_i$, at pixel level:
+For the disaggregation model, we first model the rate, $pred_i$, at pixel level:
 %
@@
 The Gaussian random field has a Matern covariance function, with hyperparameters $\rho = \frac{\sqrt{8\lambda}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating the magntiude of the variation of the field.
-Each of the model parameters and hyperparameters are given priors, which is discussed later.
+As we are working in a Bayesian setting, each of the model parameters and hyperparameters are given a prior, which is discussed later.
 

### 2019-11-21 article.Rnw
Had this problem before. But I think we want to make it clear these are likelihoods (with sim or something). But then you need to distinguish data and estimates.

@@
 
-The different likelihoods correspond to slightly different models ($y_j$ is the repsonse count data):
+The different likelihoods correspond to slightly different models ($y_j$ is the response count data):
 
@@
 \begin{equation} \label{eq:gaussian}
-N(y_j/\sum agg_i, rate_j, \sigma_j)
+y_j \sim Normal(count_j/\sum agg_i, rate_j, \sigma_j)
 \end{equation}
@@
 
-\item \textbf{Binomial} (For a survey in polygon j, $y_j$ is the number positive and $N_j$ is the number tested),
+\item \textbf{Binomial} (For a survey in polygon $j$, $y_j$ is the number positive cases and $N_j$ is the number tested),
 \begin{equation} \label{eq:binomial}
-Binom(y_j, N_j, rate_j)
+y_j \sim Binom(cases_j, N_j, rate_j)
 \end{equation}
@@
 \begin{equation} \label{eq:poisson}
-Pois(y_j, cases_j)
+y_j \sim Pois(cases_j)
 \end{equation}

### 2019-11-21 article.Rnw
Add PC refs.

@@
 
-Note the normal likelihood is modelled in rate space as it is unreasonable to assume that the dispersion of the likelihood is the same for polygons containing different numbers of cases.
+Note the normal likelihood is modelled in rate space as it is rarely reasonable to assume that the dispersion of the normal likelihood is equal for polygons containing different numbers of cases.
 Also, the binomial likelihood can be used for survey data, for example when the response data consists of the number of people who tested positive for a disease and the total number of people that were tested.
@@
 
-For each of the model parameters we specify priors and hyperpriors.
+For each of the model parameters and hyperparameters we specify priors and hyperpriors.
 The regression parameters and intercept are given Gaussian priors, where the default priors are $\beta_0 \sim N(0, 2)$ and $\beta_i \sim N(0, 0.4)$.
 
-The field parameters are given penalised complexity priors, $\rho_{min}$ and $\rho_{prob}$, where $P(\rho < \rho_{min}) = \rho_{prob}$, and the variation, $\sigma_{max}$ and $\sigma_{prob}$, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
+The field parameters are given penalised complexity priors \cite{fuglstad2018constructing}, $\rho_{min}$ and $\rho_{prob}$, where $P(\rho < \rho_{min}) = \rho_{prob}$, and the variation, $\sigma_{max}$ and $\sigma_{prob}$, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
 The effect of these hyperpriors is to regularise the field to have smaller variations over larger distances.
-Similarly for the iid effect, the priors are penalised complexity priors on the standard deviation, $sd_{max}$ and $sd_{prob}$, such that $P(sd > sd_{max}) = sd_{prob}$.
+Similarly for the iid effect, the priors are penalised complexity priors \cite{simpson2017penalising} on the standard deviation, $sd_{max}$ and $sd_{prob}$, such that $P(sd > sd_{max}) = sd_{prob}$.
 This has the effect of regularising the iid effect to be small.

### 2019-11-21 article.Rnw
Just trying to clarify that the computation is done in cpp but you, the user, work in R.

@@
 \pkg{TMB} combines the packages CppAD~\cite{bell2012cppad}, for automatic differentiation in C++, Eigen~\cite{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\cite{chen2008algorithm}, a package for efficient computation of sparse matrices.
-The use of these packages allows an efficient implementation of the applied Laplace approximation~\cite{skaug2006automatic} with exact derivatives.
+The use of these packages allows an efficient implementation of the automatic Laplace approximation~\cite{skaug2006automatic} with exact derivatives.
 \pkg{TMB} calculates first and second order derivatives of the objective function using automatic differentiation~\cite{griewank2008evaluating}.
@@
 The \pkg{TMB} package then calculates estimates of both parameters and random effects using the Laplace approximation for the likelihood.
-Evaluation of the objective function and its derivatives is performed in \proglang{R}.
+Evaluation of the objective function and its derivatives is performed via \proglang{R}.
 

### 2019-11-21 article.Rnw
clarification.

@@
 
-The \pkg{disaggregation} package is built on Template Model Builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for buidling flexible complex models based on C++.
+The \pkg{disaggregation} package is built on Template Model Builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for flexibly building complex models based on C++.
 \pkg{TMB} combines the packages CppAD~\cite{bell2012cppad}, for automatic differentiation in C++, Eigen~\cite{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\cite{chen2008algorithm}, a package for efficient computation of sparse matrices.
-The use of these packages allows an efficient implementation of the automatic Laplace approximation~\cite{skaug2006automatic} with exact derivatives.
+The use of these packages allows an efficient implementation of the automatic Laplace approximation~\cite{skaug2006automatic} with exact derivatives which gives an approximation to the Bayesian posterior.
 \pkg{TMB} calculates first and second order derivatives of the objective function using automatic differentiation~\cite{griewank2008evaluating}.

### 2019-11-21 article.Rnw
tweaks.

@@
 
-In this section we show how to use the \pkg{disaggregation} package using real malaria data of aggregated malaria case counts across Madagascar.
-Malaria is a infectious disease caused by parasites of the \emph{Plasmodium} group, transmitted by \emph{Anopheles} mosquitoes.
+In this section we show how to use the \pkg{disaggregation} package using a dataset of aggregated malaria case counts across Madagascar.
+Malaria is an infectious disease caused by parasites of the \emph{Plasmodium} group, transmitted by \emph{Anopheles} mosquitoes.
 Malaria transmission is therefore closely related to mosquito and parasite development.
@@
 Therefore, such environmental covariates would be useful in a malaria disaggregation model to inform fine-scale distributions.
-In this model we use the environmental covariates of mean land surface temperature (LST), variation in LST, elevation, and enhanced vegetation index (EVI) to inform spatial hetrogenity in malaria risk.
+In this model we use the environmental covariates of mean land surface temperature (LST), variation in LST, elevation, and enhanced vegetation index (EVI) to inform spatial heterogeneity in malaria risk.
 These covaraites are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).

### 2019-11-21 article.Rnw
Gotta keeo line length below 80 I tihnk.

@@
 
-# Make a RasterStack of covariates to the extent of Madagascar and scale values
+# Make a RasterStack of covariates, 
+# crop to the extent of Madagascar and scale values
 covariates <- lapply(covariate_raster_paths, raster)

### 2019-11-21 article.Rnw
Not sure but its weird having api in the code without some explanation.
Line length. But yes we definitely need to describe api.

@@
 These covaraites are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
+Malaria incidence rate is often given per thousand cases and given the term annual parasite index (API).
 
@@
 
-# Add incidence information to the shapefile (incidence = API * population/1000)
+# Add incidence information to the shapefile 
+# (incidence count = API * population/1000)
 indices <- match(shapes$ID_2, as.character(api$fdef_id))

### 2019-11-21 article.Rnw
Should suppress the cbind warning.

@@
 Firstly, we use the \code{prepare_data} function to setup all the data in the format needed in the disaggregation modelling.
-This function performs various data manipulation tasks to create objects that are required to provide for fitting the model.
-The required input data for the \code{prepare_data} function are a SpatialPolygonDataFrame containing the response data and a RasterStack of covariates to be used in the model.
-The variable names in the SpatialPolygonDataFrame of the response count data and the polygon ID are defined by the user.
+This function performs various data manipulation tasks to create objects that are necessary for fitting the model.
+The required input data for the \code{prepare_data} function are a SpatialPolygonsDataFrame containing the response data and a RasterStack of covariates to be used in the model.
+The variable names in the SpatialPolygonsDataFrame of the response count data and the polygon ID are defined by the user.
 
@@
 If no aggregation raster is provided a uniform distribution is assumed, i.e.
-the pixel predictions are aggregated to polygon level by summing the pixel values.
+the pixel predictions are aggregated to polygon level by summing the pixel values unaltered.
 
@@
 
-In order to know which pixels are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
+In order to know which pixels (i.e.
+which rows) are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
 Additionally, an INLA mesh is built to use for the spatial field (\code{build_mesh} function).
@@
 %
-<<prepare_data>>=
+<<prepare_data, warning = FALSE>>=
 dis_data <- prepare_data(polygon_shapefile = shapes, 

### 2019-11-21 article.Rnw
The raster plots are weird in pdf.
tweaks.

@@
 In this model we use the environmental covariates of mean land surface temperature (LST), variation in LST, elevation, and enhanced vegetation index (EVI) to inform spatial heterogeneity in malaria risk.
-These covaraites are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
+These covariates are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
 Malaria incidence rate is often given per thousand cases and given the term annual parasite index (API).
@@
 %
-<<showdata>>=
+<<showdata, dev='png'>>=
 summary(dis_data)
@@
 Now we can fit the disaggregation model using \code{fit_model}.
-Here we use a poisson likelihood for the incident count data with a log link function.
-Options for the likelihood are gaussian, poisson and binomial, and options for the link function are logit, log and identity.
+Here we use a Poisson likelihood for the incident count data with a log link function.
+Options for the likelihood are Gaussian, Poisson and binomial, and options for the link function are logit, log and identity.
 The spatial field and iid effect are components of the model by default, they can be turned off using the \code{field} and \code{iid} flags.
@@
 %
-<<fit_model>>=
+<<fit_model, message = FALSE>>=
 fitted_model <- fit_model(data = dis_data,
@@
 The \code{fit_model} function returns an object of class \code{fit.result}, which is designed to be used directly in the \code{predict} function.
-Therefore, now once we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.
+Therefore, now we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.
 
 To predict over a different spatial extent to that used in the model, a RasterStack covering the region to make predictions over can be passed as the \code{newdata} argument.
-If this argument is not given, predictions are made over the covaraite data used in the fit.
+If this argument is not given, predictions are made over the covariate rasters used in the fit.
 If the user wants to include the iid effect from the model as a component in the prediction then the \code{predict_iid} logical flag should be set to TRUE, otherwise, the iid effect will not be predicted.
@@
 The mean predictions contain a raster of the mean prediction of the incidence rate, as well as rasters of the field, iid (if predicted) and covariate component of the linear predictor.
-The uncertainty contains a RasterStack of the prediction realisations and a RasterStack of the upper and lower credible intervals.
+The \code{uncertainty} object contains a RasterStack of the prediction realisations and a RasterStack of the upper and lower credible intervals.
 
@@
 %
-<<showpredictions>>=
+<<showpredictions, dev = 'png'>>=
 plot(model_predictions$mean_predictions)

### 2019-11-21 article.Rnw
add nuts ref.

@@
 
-In this section we show performance comparisons between modelling using this \pkg{disaggregation} package, based on \pkg{TMB}, and using Markov chain Monte Carlo (MCMC).
+In this section we show performance comparisons between modelling using the Laplace approximation provided by the \pkg{disaggregation} package, based on \pkg{TMB}, and using Markov chain Monte Carlo (MCMC).
 MCMC algorithms sample the probability space to produce an esimate of the probability distribution.
@@
 The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package.
-Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm using four chains.
+Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm \cite{hoffman2014no} using four chains.
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.

### 2019-11-21 article.Rnw
If you need to run a model for 90 hours, you do it. its just annoying.

@@
 This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations, doubling the number of iterations each time until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
-The model is optimised using \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
+In contrast, fitting the model using the Laplace approximation via \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
 Fitted parameter values for both of these methods are given in Table~\ref{tab:mcmcresults}.
@@
 # Plot the trace of the parameter values for each sampling method
-stan_trace(mcmc_out, pars = c('intercept', 
+stan_trace(mcmc_out, 
+           pars = c('intercept', 
                     'slope[1]', 'slope[2]', 'slope[3]', 'slope[4]', 
@@
 It can be seen that the MCMC algorithm has been run for long enough to get sufficient chain mixing.
-The \pkg{disaggregation} package produces similar results to the MCMC algorithm, however the model fitting using the \pkg{disaggregation} package only takes 56 seconds.
-Therefore, it can be seen that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that cannot practically be done using MCMC.
+The \pkg{disaggregation} package produces similar results to the MCMC algorithm.
+However the model fitting using the \pkg{disaggregation} package took 56 seconds in contrast to the 94 hours taken for the MCMC run.
+Therefore, it can be seen that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that can be prohibitively slow using MCMC.
 

### 2019-11-22 article.Rnw

@@
 
-Disaggregation modelling, which involves predicting at fine-scale resolution using areal data over heterogenous regions, has become widely used in fields such as epidemiology and ecology.
+Disaggregation modelling, which involves fitting models at fine-scale resolution using areal data over heterogenous regions, has become widely used in fields such as epidemiology and ecology.
 The \pkg{disaggregation} package implements Bayesian spatial disaggregation modelling with a simple, easy to use \proglang{R} interface.
@@
 These disaggregation models are computationally intensive and take a long time using Markov Chian Monte Carlo (MCMC) optimisation techniques.
-Using \pkg{TMB}, the models are much faster and produce similar results
+Using \pkg{TMB}, the models are much faster and produce similar results.
 
-The \pkg{disaggregation} package can be extended to include spatio-temporal disaggregation models.
+Future work will entail extending the \pkg{disaggregation} package to include spatio-temporal disaggregation models.
 This would require a spatio-temporal field as well as dynamic covariates.

### 2019-11-22 article.Rnw
Correct typo in table

@@
 $log(\sigma)$ & -3.34 & 0.59 & -3.55 & 0.59 \\
-$log(\rho)$ & 0.72 & 0.31 & -0.56 & 0.30 \\
+$log(\rho)$ & 0.72 & 0.31 & 0.56 & 0.30 \\
 \end{tabular}

### 2019-11-29 article.Rnw
Typo count to cases
Use cropped prepared data for direct use in the package

@@
 \begin{equation} \label{eq:gaussian}
-y_j \sim Normal(count_j/\sum agg_i, rate_j, \sigma_j)
+y_j \sim Normal(cases_j/\sum agg_i, rate_j, \sigma_j)
 \end{equation}
@@
 
-We then setup the data to use in the \pkg{disaggregation} package.
+We then read in the data to use in the \pkg{disaggregation} package.
 
@@
 
-API_path <- 'data/api_subnational.csv'
-population_path <-'data/pop_at_risk_pf_2013.tif'
-shapefile_path <- 'data/admin2019_2.shp'
-
-covariate_raster_paths <- c(
-  'data/MODIS_LST_Day.mean.5km.tif',
-  'data/MODIS_EVI.mean.5km.tif',
-  'data/SRTM_elevation.5km.mean.tif',
-  'data/MODIS_LST_Day.SD.5km.tif')
-
-# Make a RasterStack of covariates, 
-# crop to the extent of Madagascar and scale values
-covariates <- lapply(covariate_raster_paths, raster)
-covariate_stack <- stack(covariates)
-covariate_stack <- scale(crop(covariate_stack, extent(c(42, 52, -26, -10))))
-
-# Read in the incidence data and only select data from Madagascar 2013
-api <- read.csv(API_path)
-api <- api[api$iso3 == 'MDG' & api$year == 2013, ]
-
-# Read in the shapefile and only keep those that contain incidence data
-shapes <- shapefile(shapefile_path)
-shapes <- shapes[shapes$ID_2 %in% as.character(api$fdef_id), ]
-
-# Add incidence information to the shapefile 
-# (incidence count = API * population/1000)
-indices <- match(shapes$ID_2, as.character(api$fdef_id))
-shapes$inc <- (api$api_mean_pf[indices] * api$pop_at_risk_pf[indices])/1000
-
-# Read in the population raster and crop it to the extent of Madagascar
-population_raster <- raster(population_path)
-population_raster <- crop(population_raster, extent(c(42, 52, -26, -10)))
+shapes <- shapefile('data/shapes/mdg_shapes.shp')
+population_raster <- raster('data/population.tif')
+covariate_stack <- getCovariateRasters('data/covariates', shape = population_raster)
 @

### 2019-12-02 article.Rnw
Adding Rohan's edits

@@
 \newcommand{\fct}[1]{\code{#1()}}
+\usepackage{amsmath}
 
@@
 
-Suppose we have response data, $y_j$, for $N$ polygons, which corresponds to count data for the property of interest within that polygon. 
+Suppose we have response data, $y_i$, for $N$ polygons, which corresponds to count data for the property of interest within that polygon. 
 The process that is being counted occurs in continuous space that we model as a high-resolution, square lattice for convenience.
-The data, $y_j$,  are assumed to be created by the aggregation of the counts over the polygon, i.e.
+The data, $y_i$,  are assumed to be created by the aggregation of the counts over the polygon, i.e.
 the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
-The rate, $p_i$, is defined such that the number of cases in a pixel can be calculated by multiplying the rate by the aggregation raster.
+The rate is defined such that the number of cases in this pixel can be calculated by multiplying the rate by the aggregation raster.
 For example, in epidemiology, we may have as our response the number of people that contract a certain disease in a given period of time (case incidence).
@@
 
-For the disaggregation model, we first model the rate, $pred_i$, at pixel level:
+For the disaggregation model, we first model the rate at pixel level, with the rate in pixel $j$ of polygon $i$ given by:
 %
 \begin{equation} \label{eq:lp}
-link(pred_i) = \beta_0 + \beta X + GP(s_i) + u_j
+\textrm{link}(\textrm{rate}_{ij}) = \beta_0 + \beta X_{ij} + \textrm{GP}(s_{ij}) + u_i 
 \end{equation}
 %
-
-where $\beta$ are the regression coefficients, $X$ are the covariates, $GP$ is the gaussian random field and $u_j$ is the polygon-specific iid random effect.
-The user defined link function is given by $link$.
-The Gaussian random field has a Matern covariance function, with hyperparameters $\rho = \frac{\sqrt{8\lambda}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating the magntiude of the variation of the field.
+Here $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific effect.
+The user defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively  The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
 As we are working in a Bayesian setting, each of the model parameters and hyperparameters are given a prior, which is discussed later.
 
-The pixel predictions are then aggregated to the polygon level using the weighted sum (via the aggregation raster, $agg_i$):
+The pixel predictions are then aggregated to the polygon level using the weighted sum (via the aggregation raster $a_{ij}$):
 
@@
 \begin{equation} \label{eq:agreggate1}
-cases_j = \sum_{i \epsilon j} pred_i \times agg_i
+\textrm{cases}_i = \sum_{j=1}^{N_i} a_{ij}\textrm{rate}_{ij}
 \end{equation}
 \begin{equation} \label{eq:agreggate2}
-rate_j = \frac{\sum_{i \epsilon j} pred_i \times agg_i}{\sum_{i \epsilon j} agg_i}
+\textrm{rate}_i = \frac{\textrm{cases}_{i}}{\sum_{j=1}^{N_i} a_{ij}}
 \end{equation}
 %
-
-The different likelihoods correspond to slightly different models ($y_j$ is the response count data):
+The different likelihoods correspond to slightly different models ($y_i$ is the response count data):
 
@@
 \begin{itemize}
-\item {\textbf{Gaussian} ($\sigma_j$ is the dispersion of the polygon data)
+\item {\textbf{Gaussian} ($\sigma_i$ is the dispersion of the polygon data)
 \begin{equation} \label{eq:gaussian}
-y_j \sim Normal(cases_j/\sum agg_i, rate_j, \sigma_j)
+y_i \sim \textrm{Normal}(\textrm{cases}_i, \sigma_i)
 \end{equation}
-Here $\sigma_j = \sigma \sqrt{\sum agg_i^2} / \sum agg_i$, where  $\sigma$ is the dispersion of the pixel data (a parameter learnt by the model) }
+Here $\sigma_i = \sigma \sqrt{\sum_j a_{ij}^2}$, where  $\sigma$ is the pixel-level dispersion (a parameter learnt by the model) }
 
-\item \textbf{Binomial} (For a survey in polygon $j$, $y_j$ is the number positive cases and $N_j$ is the number tested),
+\item \textbf{Binomial} (For a survey in polygon $i$, $y_i$ is the number positive cases and $M_i$ is the number tested),
 \begin{equation} \label{eq:binomial}
-y_j \sim Binom(cases_j, N_j, rate_j)
+y_i \sim \textrm{Binomial}(M_i, \textrm{rate}_i)
 \end{equation}
@@
 \begin{equation} \label{eq:poisson}
-y_j \sim Pois(cases_j)
+y_i \sim \textrm{Pois}(\textrm{cases}_i)
 \end{equation}
@@
 
-For each of the model parameters and hyperparameters we specify priors and hyperpriors.
+For each of the model parameters and hyperparameters we specify priors.
 The regression parameters and intercept are given Gaussian priors, where the default priors are $\beta_0 \sim N(0, 2)$ and $\beta_i \sim N(0, 0.4)$.
-
-The field parameters are given penalised complexity priors \cite{fuglstad2018constructing}, $\rho_{min}$ and $\rho_{prob}$, where $P(\rho < \rho_{min}) = \rho_{prob}$, and the variation, $\sigma_{max}$ and $\sigma_{prob}$, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
-The effect of these hyperpriors is to regularise the field to have smaller variations over larger distances.
-Similarly for the iid effect, the priors are penalised complexity priors \cite{simpson2017penalising} on the standard deviation, $sd_{max}$ and $sd_{prob}$, such that $P(sd > sd_{max}) = sd_{prob}$.
-This has the effect of regularising the iid effect to be small.
+A penalised complexity prior is placed on the scale and range parameters of the random field \cite{fuglstad2018constructing} such that
+\begin{align}
+    P(\rho < \rho_{min}) &= \rho_{prob} \\
+    P(\sigma > \sigma_{max}) &= \sigma_{prob}
+\end{align}
+where the values $\rho_{min}, \,\rho_{prob}, \,\sigma_{max}, \,\sigma_{prob}$ are set by the user.
+This prior shrinks the field towards a base model with zero variance and infinite range, in other words regularising towards a flatter field with smaller magnitude.
+The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $u_{sd}$.
+A penalised complexity prior \cite{simpson2017penalising} such that $P(u_{sd} > sd_{max}) = sd_{prob}$ which shrinks towards a base model of no polygon-specific effect.
+The $sd_{max}$ and $sd_{prob}$ are set by the user.
 
@@
 In this section we show performance comparisons between modelling using the Laplace approximation provided by the \pkg{disaggregation} package, based on \pkg{TMB}, and using Markov chain Monte Carlo (MCMC).
-MCMC algorithms sample the probability space to produce an esimate of the probability distribution.
-These algorithms tend to be slow, as it takes a long time to effectively sample the likelihood space.
-In contrast, the \pkg{disaggregation} package, which uses the Laplace approximation, automatic differentiation and efficient computation of sparse matrices within \pkg{TMB}, takes a considerably shorter time to optimise the model.
+Given a function to evaluate the probability density of a distribution at any given point in parameter space, MCMC algorithms construct markov chains to generate samples from this distribution.
+These algorithms are often slow, particularly in high-dimensional settings, as it can take a long time to converge to and effectively sample from the stationary distribution.
+The density is evaluated at each step, so this problem is compounded when evaluating this density is computationally expensive.
+In contrast, the \pkg{disaggregation} package approximates uses a Laplace approximation to the posterior to generate posterior samples.
+This only requires the posterior to be maximised to find the posterior mode and therefore involves relatively few evaluations of the posterior density compared to MCMC techniques, although potentially at the expense of less accurate posterior samples.
 Here we compare the time and performance of the two techniques.

### 2019-12-02 article.Rnw
Minor edits to likelihood text

@@
 \begin{itemize}
-\item {\textbf{Gaussian} ($\sigma_i$ is the dispersion of the polygon data)
+\item {\textbf{Gaussian}
 \begin{equation} \label{eq:gaussian}
@@
 
-\item \textbf{Binomial} (For a survey in polygon $i$, $y_i$ is the number positive cases and $M_i$ is the number tested),
+\item \textbf{Binomial}
 \begin{equation} \label{eq:binomial}
@@
 \end{equation}
+For a survey in polygon $i$, $y_i$ is the number positive cases and $M_i$ is the number tested.
 
-\item \textbf{Poisson} (predicts incidence count),
+\item \textbf{Poisson}
 \begin{equation} \label{eq:poisson}

### 2019-12-02 article.Rnw
Make iid effect clearer

@@
 This prior shrinks the field towards a base model with zero variance and infinite range, in other words regularising towards a flatter field with smaller magnitude.
-The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $u_{sd}$.
-A penalised complexity prior \cite{simpson2017penalising} such that $P(u_{sd} > sd_{max}) = sd_{prob}$ which shrinks towards a base model of no polygon-specific effect.
-The $sd_{max}$ and $sd_{prob}$ are set by the user.
+The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $\sigma_{u}$ (where the precision $\tau_{u} = 1/\sigma_{u}^2$).
+A penalised complexity prior \cite{simpson2017penalising} such that
+\begin{equation}
+P(\sigma_{u} > \sigma_{u, max}) = \sigma_{u, prob}
+\end{equation}
+which shrinks towards a base model of no polygon-specific effect. The $\sigma_{u, max}$ and $\sigma_{u, prob}$ are set by the user.
 
@@
 Slope 4 & -0.24 & 0.23 & -0.24 & 0.21 \\
-iid $log(\tau)$ & 1.08 & 0.27 & 1.10 & 0.27 \\
+$log(\tau_u)$ & 1.08 & 0.27 & 1.10 & 0.27 \\
 $log(\sigma)$ & -3.34 & 0.59 & -3.55 & 0.59 \\

### 2019-12-03 article.Rnw
Tidy edit
Tidy up plot formatting and add authors

@@
   cache = TRUE,
-  fig.width = 7
+  fig.width = 6,
+  fig.height = 4
 )
@@
    \And Tim CD Lucas\\University of Oxford
-   \And Rohan Arambepola\\University of Oxford}
-   %\AND Andre Python\\University of Oxford
-   %\And Daniel J Weiss\\University of Oxford}
+   \And Rohan Arambepola\\University of Oxford
+   \AND Peter Gething\\Telethon Kids Institute \\ Curtin University
+   \And Daniel J Weiss\\University of Oxford}
 \Plainauthor{Anita K Nandi, Tim CD Lucas, Rohan Arambepola, Andre Python, Daniel J Weiss}
@@
 <<installview, eval=FALSE>>=
-devtools::install_github("aknandi/disaggregation")
+devtools::install.packages("disaggregation")
 library(disaggregation)
@@
 %
-<<setuppaths>>=
+<<setuppaths, results='hide'>>=
 library(raster)
@@
 population_raster <- raster('data/population.tif')
-covariate_stack <- getCovariateRasters('data/covariates', shape = population_raster)
+covariate_stack <- getCovariateRasters('data/covariates', 
+                                       shape = population_raster)
 @
@@
 %
-<<showdata, dev='png'>>=
+<<summarydata>>=
 summary(dis_data)
+@
+%
+
+%
+<<plotdata, eval=FALSE>>=
 plot(dis_data)
+@
+%
+
+%
+<<plotdatahidden, echo=FALSE, results='hide', fig.show='hide'>>=
+data_plots <- plot(dis_data)
+@
+%
+
+%
+<<plotdata1, echo=FALSE, dev='png', fig.width = 8, fig.height = 6>>=
+data_plots$polygon
+@
+%
+
+%
+<<plotdata2, echo=FALSE, dev='png', fig.width = 8, fig.height = 5>>=
+data_plots$covariates
+@
+%
+
+%
+<<plotdata3, echo=FALSE, dev='png', fig.width = 5, fig.height = 5>>=
+INLA::plot.inla.mesh(dis_data$mesh, main = 'INLA mesh for spatial field')
 @
@@
 %
-<<showmodel>>=
+<<summarymodel>>=
 summary(fitted_model)
+@
+%
+
+%
+<<plotmodel, fig.width = 10, fig.height = 6>>=
 plot(fitted_model)
@@
 %
-<<showpredictions, dev = 'png'>>=
+<<plotmean, dev = 'png', fig.width = 10, fig.height = 7>>=
 plot(model_predictions$mean_predictions)
+@
+%
+
+%
+<<plotuncertainty, dev = 'png', fig.width = 10, fig.height = 5>>=
 plot(model_predictions$uncertainty_predictions)
 @
+%
 

### 2019-12-03 article.Rnw
Correct authors
Edit intro, add ref

@@
    \And Daniel J Weiss\\University of Oxford}
-\Plainauthor{Anita K Nandi, Tim CD Lucas, Rohan Arambepola, Andre Python, Daniel J Weiss}
+\Plainauthor{Anita K Nandi, Tim CD Lucas, Rohan Arambepola, Peter Gething, Daniel J Weiss}
 
@@
 Disaggregation regression, first applied in species distribution modelling in ecology~\cite{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
-The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions.
-However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the agreggated area.
+The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions, as relationships learned between variables at one spatial scale may not hold at other scales \cite{wakefield2006health}.
+However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the aggregated area.
 

### 2019-12-03 article.Rnw
Edits to the model section

@@
 
-For the disaggregation model, we first model the rate at pixel level, with the rate in pixel $j$ of polygon $i$ given by:
+For the disaggregation model we model the rate at pixel level, with the likelihood for then observed data then given by aggregating these pixel level rates.
+The rate in pixel $j$ of polygon $i$ at location $s_{ij}$ is given by:
 %
@@
 %
-Here $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific effect.
+where $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific effect.
 The user defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively  The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
@@
 \begin{itemize}
+\item \textbf{Poisson}
+\begin{equation} \label{eq:poisson}
+y_i \sim \textrm{Pois}(\textrm{cases}_i)
+\end{equation}
+
 \item {\textbf{Gaussian}
@@
 \end{equation}
-For a survey in polygon $i$, $y_i$ is the number positive cases and $M_i$ is the number tested.
-
-\item \textbf{Poisson}
-\begin{equation} \label{eq:poisson}
-y_i \sim \textrm{Pois}(\textrm{cases}_i)
-\end{equation}
 
@@
 %
-
-Note the normal likelihood is modelled in rate space as it is rarely reasonable to assume that the dispersion of the normal likelihood is equal for polygons containing different numbers of cases.
-Also, the binomial likelihood can be used for survey data, for example when the response data consists of the number of people who tested positive for a disease and the total number of people that were tested.
+In the example of disease mapping, Poisson or Gaussian likelihoods could be used when the quantity observed, $y_i$, is the total number of cases in a given polygon.
+The Binomial model could be used when $y_i$ is the prevalence of a disease in a sample of $M_i$ people in the polygon.
 

### 2019-12-03 article.Rnw
Non-italised shape and rate

@@
 
-For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(shape = 1, rate = 5e-05)$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values.
+For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(\text{shape} = 1, \text{rate} = 5e-05)$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values.
 This is the same as the prior set by INLA for the dispersion of the normal likelihood.

### 2019-12-03 article.Rnw
Minor edit
Formatting for JSS. Refs in title case (used perl script)

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{disaggregation: An \proglang{R} package for Bayesian spatial disaggregation modelling}
-\Plaintitle{disaggregation: An R package for Bayesian spatial disaggregation modelling}
-\Shorttitle{disaggregation modelling in \proglang{R}}
+\title{disaggregation: An \proglang{R} Package for Bayesian Spatial Disaggregation Modelling}
+\Plaintitle{disaggregation: An R Package for Bayesian Spatial Disaggregation Modelling}
+\Shorttitle{Disaggregation Modelling in \proglang{R}}
 
@@
 the number of rows in the covariate data is different to that of the response data.
-The spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate}, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \cite{wilson2017pointless}.
-Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
+The spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate} (integrated nested Laplace approximation), has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \cite{wilson2017pointless}.
+Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} (template model builder) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
 The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models at the expense of some flexibility.
@@
 
-The \pkg{disaggregation} package is built on Template Model Builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for flexibly building complex models based on C++.
+The \pkg{disaggregation} package is built on template model builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for flexibly building complex models based on C++.
 \pkg{TMB} combines the packages CppAD~\cite{bell2012cppad}, for automatic differentiation in C++, Eigen~\cite{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\cite{chen2008algorithm}, a package for efficient computation of sparse matrices.
@@
 %
-<<setuppaths, results='hide'>>=
+<<setuppaths, results='hide', show='hide'>>=
 library(raster)
@@
                                           offset = c(1, 2)),
-                         id_var ='ID_2',
+                         id_var = 'ID_2',
                          response_var = 'inc',
@@
 
-\section{Comparison with MCMC} \label{sec:comparison}
+\section{Comparison with Markov chain Monte Carlo (MCMC)} \label{sec:comparison}
 
-In this section we show performance comparisons between modelling using the Laplace approximation provided by the \pkg{disaggregation} package, based on \pkg{TMB}, and using Markov chain Monte Carlo (MCMC).
+In this section we show performance comparisons between modelling using the Laplace approximation provided by the \pkg{disaggregation} package, based on \pkg{TMB}, and using MCMC.
 Given a function to evaluate the probability density of a distribution at any given point in parameter space, MCMC algorithms construct markov chains to generate samples from this distribution.
@@
 The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package.
-Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS algorithm \cite{hoffman2014no} using four chains.
+Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS (no-u-turn sampler) algorithm \cite{hoffman2014no} using four chains.
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
@@
 This allows fast, optimised calculations in C++.
-These disaggregation models are computationally intensive and take a long time using Markov Chian Monte Carlo (MCMC) optimisation techniques.
+These disaggregation models are computationally intensive and take a long time using MCMC optimisation techniques.
 Using \pkg{TMB}, the models are much faster and produce similar results.

### 2019-12-03 article.Rnw
minor change

@@
 
-For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(\text{shape} = 1, \text{rate} = 5e-05)$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values.
+For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(\text{shape} = 1, \text{rate} = 5 \times 10^{-5})$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values.
 This is the same as the prior set by INLA for the dispersion of the normal likelihood.

### 2019-12-04 article.Rnw
Formatting figures. Captions and figure references. Other minor changes

@@
 \usepackage{amsmath}
+\usepackage{subfig}
+\usepackage{float}
 
@@
   fig.width = 6,
-  fig.height = 4
+  fig.height = 4,
+  fig.pos = 'H'
 )
@@
 %
-where $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific effect.
+where $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific iid effect.
 The user defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively  The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
@@
 Additionally, an INLA mesh is built to use for the spatial field (\code{build_mesh} function).
-The \code{mesh.args} argument allows you to supply a list of INLA mesh parameters to control the mesh used for the spatial field.
-The mesh can take a while to construct, so if you want to prepare the data without buidling the mesh, the user can set the \code{makeMesh} flag to FALSE.
-However, you cannot then fit the disaggregation model without the mesh.
+The \code{mesh.args} argument allows the user to supply a list of INLA mesh parameters to control the mesh used for the spatial field.
+The mesh can take a while to construct, so to prepare the data without buidling the mesh the user can set the \code{makeMesh} flag to FALSE.
+However, it would then not be possible to fit the disaggregation model without the mesh.
 
@@
 The summary function returns information on how many pixels and polygons the data contains, how many pixels in the smallest and largest polygons and a summary of the covariate data.
-The \code{plot} functions plots a map of the polygon response data, the covariate rasters and the INLA mesh.
+The \code{plot} functions plots a map of the polygon response data, the covariate rasters and the INLA mesh, as shown in Figure~\ref{fig:subplotdata}.
 
@@
 %
-<<plotdata1, echo=FALSE, dev='png', fig.width = 8, fig.height = 6>>=
+<<subplotdata, fig.cap='Maps of Madagascar showing the data used in the disaggregation model.
+These plots are produced when calling the plot function on the disag.data object.', fig.subcap=c('Number of malaria cases in each administrative unit', 'Maps of the four covariates used in the model', 'inla.mesh object that will be used to make the spatial field'), out.width='0.5\\linewidth', echo=FALSE, fig.ncol = 2, dev = 'png', fig.width = 4, fig.height = 5, fig.pos = 'H'>>=
 data_plots$polygon
-@
-%
-
-%
-<<plotdata2, echo=FALSE, dev='png', fig.width = 8, fig.height = 5>>=
 data_plots$covariates
-@
-%
-
-%
-<<plotdata3, echo=FALSE, dev='png', fig.width = 5, fig.height = 5>>=
 INLA::plot.inla.mesh(dis_data$mesh, main = 'INLA mesh for spatial field')
@@
 Hyperpriors for the field are given as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{min}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{min}) = \sigma_{prob}$.
-Similarly, you specify penalised complexity priors for the iid effect.
+Similarly, the user specifies penalised complexity priors for the iid effect.
 The \code{iterations} argument specifies the maximum number of iterations the model can run for to find an optimal point.
-In order to print more verbose output you can set the \code{silent} argument to FALSE.
+In order to print more verbose output the user can set the \code{silent} argument to FALSE.
 
@@
 The \code{summary} function gives the estimate and standard error of the fixed effect parameters in the model, as well as the negative log likelihood and in-sample performance metrics (root mean squared error, mean absolute error, pearson correlation coefficient and spearman rank correlation coefficient).
-The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions.
+The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions, as shown in Figure~\ref{fig:fitmodel}.
 
@@
 
+\begin{figure}[!h]
+\centering
 %
@@
 %
+\caption{\label{fig:fitmodel} Plot summarising the results of the fitted model.
+These plots are produced when calling the \code{plot} function on the \code{fit.result} object.
+The fixed effects plot (left) shows the fitted parameter values with uncertainty estimation for all the fixed effects in the model.
+The in-sample performance plot (right) shows the predicted incidence rate values for each polygon in the data against the observed values for that polygon in the data.}
+\end{figure}
 
@@
 
-The \code{plot} function can be used on both the \code{predictions} and \code{uncertainty} objects.
+The \code{plot} function can be used on both the \code{predictions} and \code{uncertainty} objects, as shown in Figures~\ref{fig:subpredictmean} and \ref{fig:predictuncertainty} respectively.
 
 %
-<<plotmean, dev = 'png', fig.width = 10, fig.height = 7>>=
+<<plotpredictions, eval=FALSE>>=
 plot(model_predictions$mean_predictions)
@@
 %
+<<hiddenplotmean, echo=FALSE, results='hide', fig.show='hide'>>=
+mean_predictions <- plot(model_predictions$mean_predictions)
+@
+%
+
+%
+<<subpredictmean, fig.cap='Maps of Madagascar showing the fine-scale predictions from the disaggregation model.
+These maps are produced when calling the plot function on the predictions object.
+The maps show the (a) mean prediction of malaria incidence rate, (b) covariate contribution to the linear predictor, and (c) field contribution to the linear predictor', fig.subcap=c('', '', ''), out.width='0.33\\linewidth', echo=FALSE, dev = 'png', fig.width = 4, fig.height = 6, fig.pos = 'H'>>=
+mean_predictions$mean
+mean_predictions$covariates
+mean_predictions$field
+@
+%
+
+\begin{figure}
+\centering
+%
 <<plotuncertainty, dev = 'png', fig.width = 10, fig.height = 5>>=
@@
 %
+\caption{\label{fig:predictuncertainty} Maps of Madagascar showing the fine-scale predictions of lower (2.5\%) and upper (97.5\%) credible intervals of the malaria incidence rate from the disaggregation model.
+These maps are produced when calling the \code{plot} function on the \code{uncertainty} object.}
+\end{figure}
 
-Using three simple functions we have been able to fit a Bayesian spatial disaggregation model and predict pixel-level incidence rate across Madagascar using aggregated incidence data and pixel-level environmental covariates.
+Using three simple functions within the \pkg{disaggregation} package we have been able to fit a Bayesian spatial disaggregation model and predict pixel-level incidence rate across Madagascar using aggregated incidence data and pixel-level environmental covariates.
 
+\newpage
 \section{Comparison with Markov chain Monte Carlo (MCMC)} \label{sec:comparison}
@@
 \includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 94 hours.}
+\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 94 hours. It was run for 8000 iterations with 2000 of those as warmup.}
 \label{fig:mcmclong}
@@
 
-Future work will entail extending the \pkg{disaggregation} package to include spatio-temporal disaggregation models.
-This would require a spatio-temporal field as well as dynamic covariates.
-However, this would be significantly more computationally intensive.
+Future work could be done extending the \pkg{disaggregation} package to include spatio-temporal disaggregation models.
+This would require a spatio-temporal field as well as dynamic covariates, and would be significantly more computationally intensive.
 Additionally, tools for cross-validation could be included within the package.

### 2019-12-10 article.Rnw
Dan's comments

@@
 
-For the disaggregation model we model the rate at pixel level, with the likelihood for then observed data then given by aggregating these pixel level rates.
+For the disaggregation model we model the rate at pixel level, with the likelihood for the observed data given by aggregating these pixel level rates.
 The rate in pixel $j$ of polygon $i$ at location $s_{ij}$ is given by:
@@
 where $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific iid effect.
-The user defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively  The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
+The user-defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively.
+The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
 As we are working in a Bayesian setting, each of the model parameters and hyperparameters are given a prior, which is discussed later.
@@
 The \code{mesh.args} argument allows the user to supply a list of INLA mesh parameters to control the mesh used for the spatial field.
-The mesh can take a while to construct, so to prepare the data without buidling the mesh the user can set the \code{makeMesh} flag to FALSE.
+The mesh can take several minutes to construct, so to prepare the data without buidling the mesh the user can set the \code{makeMesh} flag to FALSE.
 However, it would then not be possible to fit the disaggregation model without the mesh.
@@
 Ideally the NAs in the data would be dealt with by the user beforehand, however, setting na.action = TRUE will automatically deal with NAs.
-It removes any polygons that have NAs as a response, sets any aggregation pixels with NA to zero and sets covariate NAs pixels to the median value for the that covariate.
+It removes any polygons that have NAs as a response, sets any aggregation pixels with NA to zero and sets covariate NAs pixels to the median value for that covariate across all polygons.
 
@@
 <<subplotdata, fig.cap='Maps of Madagascar showing the data used in the disaggregation model.
-These plots are produced when calling the plot function on the disag.data object.', fig.subcap=c('Number of malaria cases in each administrative unit', 'Maps of the four covariates used in the model', 'inla.mesh object that will be used to make the spatial field'), out.width='0.5\\linewidth', echo=FALSE, fig.ncol = 2, dev = 'png', fig.width = 4, fig.height = 5, fig.pos = 'H'>>=
+These plots are produced when calling the plot function on the disag.data object.
+The plots show: (a) number of malaria cases in each administrative unit, (b) maps of the four covariates used in the model; and (c) inla.mesh object that will be used to make the spatial field.', fig.subcap=c('', '', ''), out.width='0.5\\linewidth', echo=FALSE, fig.ncol = 2, dev = 'png', fig.width = 4, fig.height = 5, fig.pos = 'H'>>=
 data_plots$polygon
@@
 %
-<<plotmodel, fig.width = 10, fig.height = 6>>=
+<<plotmodel, fig.width = 13, fig.height = 6>>=
 plot(fitted_model)
@@
 The \code{fit_model} function returns an object of class \code{fit.result}, which is designed to be used directly in the \code{predict} function.
-Therefore, now we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.
+Therefore, now that we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.
 
@@
 These maps are produced when calling the plot function on the predictions object.
-The maps show the (a) mean prediction of malaria incidence rate, (b) covariate contribution to the linear predictor, and (c) field contribution to the linear predictor', fig.subcap=c('', '', ''), out.width='0.33\\linewidth', echo=FALSE, dev = 'png', fig.width = 4, fig.height = 6, fig.pos = 'H'>>=
+The maps show the (a) mean prediction of malaria incidence rate, (b) covariate contribution to the linear predictor, and (c) field contribution to the linear predictor.', fig.subcap=c('', '', ''), out.width='0.33\\linewidth', echo=FALSE, dev = 'png', fig.width = 4, fig.height = 6, fig.pos = 'H'>>=
 mean_predictions$mean
@@
 The \pkg{disaggregation} package implements Bayesian spatial disaggregation modelling with a simple, easy to use \proglang{R} interface.
-The package includes simple data preparation, fitting and prediction functions that allow some user defined model flexibility.
+The package includes simple data preparation, fitting and prediction functions that allow some user-defined model flexibility.
 In this paper we have presented an application of the package, predicting malaria incidence rate across Madagascar from aggregated count data and environmental covariates.

### 2019-12-10 article.Rnw
Remove raster package loading message

@@
 library(disaggregation, quietly = TRUE)
+library(raster, quietly = TRUE)
 @

### 2019-12-10 article.Rnw
Easier to read citations

@@
 Methods for estimating high-resolution risk maps from aggregated response data over large spatial regions are 
-becoming increasingly sought after in disease risk mapping~\cite{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\cite{sturrock2014fine, sturrock2016mapping}. 
-Disaggregation regression, first applied in species distribution modelling in ecology~\cite{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \cite{falciparum2019, vivax2019}.
-The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions, as relationships learned between variables at one spatial scale may not hold at other scales \cite{wakefield2006health}.
+becoming increasingly sought after in disease risk mapping~\citep{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\citep{sturrock2014fine, sturrock2016mapping}. 
+Disaggregation regression, first applied in species distribution modelling in ecology~\citep{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \citep{falciparum2019, vivax2019}.
+The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions, as relationships learned between variables at one spatial scale may not hold at other scales \citep{wakefield2006health}.
 However, by using fine-scale information from related covariates we can inform the hetrogeneity of the response variable of interest within the aggregated area.
@@
 the number of rows in the covariate data is different to that of the response data.
-The spatial modelling software package, \pkg{INLA}~\cite{rue2009approximate} (integrated nested Laplace approximation), has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \cite{wilson2017pointless}.
-Disaggregation models can be implemented in \pkg{TMB}~\cite{tmb2016} (template model builder) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
+The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate} (integrated nested Laplace approximation), has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
+Disaggregation models can be implemented in \pkg{TMB}~\citep{tmb2016} (template model builder) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
 The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models at the expense of some flexibility.
@@
 The regression parameters and intercept are given Gaussian priors, where the default priors are $\beta_0 \sim N(0, 2)$ and $\beta_i \sim N(0, 0.4)$.
-A penalised complexity prior is placed on the scale and range parameters of the random field \cite{fuglstad2018constructing} such that
+A penalised complexity prior is placed on the scale and range parameters of the random field \citep{fuglstad2018constructing} such that
 \begin{align}
@@
 The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $\sigma_{u}$ (where the precision $\tau_{u} = 1/\sigma_{u}^2$).
-A penalised complexity prior \cite{simpson2017penalising} such that
+A penalised complexity prior \citep{simpson2017penalising} such that
 \begin{equation}
@@
 The \pkg{disaggregation} package is built on template model builder (\pkg{TMB})~\cite{tmb2016}, which is a tool for flexibly building complex models based on C++.
-\pkg{TMB} combines the packages CppAD~\cite{bell2012cppad}, for automatic differentiation in C++, Eigen~\cite{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\cite{chen2008algorithm}, a package for efficient computation of sparse matrices.
-The use of these packages allows an efficient implementation of the automatic Laplace approximation~\cite{skaug2006automatic} with exact derivatives which gives an approximation to the Bayesian posterior.
-\pkg{TMB} calculates first and second order derivatives of the objective function using automatic differentiation~\cite{griewank2008evaluating}.
+\pkg{TMB} combines the packages CppAD~\citep{bell2012cppad}, for automatic differentiation in C++, Eigen~\citep{guennebaud2010eigen}, a C++ library for linear algebra, and CHOLMOD~\citep{chen2008algorithm}, a package for efficient computation of sparse matrices.
+The use of these packages allows an efficient implementation of the automatic Laplace approximation~\citep{skaug2006automatic} with exact derivatives which gives an approximation to the Bayesian posterior.
+\pkg{TMB} calculates first and second order derivatives of the objective function using automatic differentiation~\citep{griewank2008evaluating}.
 
@@
 The model described in Section~\ref{sec:code} for malaria in Madagascar has been optimised using the \pkg{disaggregation} package.
-Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS (no-u-turn sampler) algorithm \cite{hoffman2014no} using four chains.
+Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS (no-u-turn sampler) algorithm \citep{hoffman2014no} using four chains.
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.

### 2019-12-11 article.Rnw
Small changes before submission

@@
 the number of rows in the covariate data is different to that of the response data.
-The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate} (integrated nested Laplace approximation), has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
-Disaggregation models can be implemented in \pkg{TMB}~\citep{tmb2016} (template model builder) with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
+The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
+Disaggregation models can be implemented in \pkg{TMB}~\citep{tmb2016}, or template model builder, with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
 The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models at the expense of some flexibility.
@@
 The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $\sigma_{u}$ (where the precision $\tau_{u} = 1/\sigma_{u}^2$).
-A penalised complexity prior \citep{simpson2017penalising} such that
+A penalised complexity prior \citep{simpson2017penalising} is placed on $\sigma_{u}$ such that
 \begin{equation}
@@
 For the uncertainty calculations, parameter values are sampled from the posterior distribution and summarised.
-The number of parameter draws used to calculate the uncertainty is set by the user via the \code{N} parameter (default: 100), and the size of the credible interval (75\%, 95\%, ...) to be calculated when summarising is set via the argument \code{CI} (default: 0.95).
+The number of parameter draws used to calculate the uncertainty is set by the user via the \code{N} parameter (default: 100), and the size of the credible interval (e.g.
+75\%, 95\%) to be calculated when summarising is set via the argument \code{CI} (default: 0.95).
 
@@
 The density is evaluated at each step, so this problem is compounded when evaluating this density is computationally expensive.
-In contrast, the \pkg{disaggregation} package approximates uses a Laplace approximation to the posterior to generate posterior samples.
+In contrast, the \pkg{disaggregation} package approximates using a Laplace approximation to the posterior to generate posterior samples.
 This only requires the posterior to be maximised to find the posterior mode and therefore involves relatively few evaluations of the posterior density compared to MCMC techniques, although potentially at the expense of less accurate posterior samples.
@@
 
-m <- fit_model(data = dis_data, 
+tmb_model <- fit_model(data = dis_data, 
                        iterations = 1,
@@
 start <- Sys.time()
-mcmc_out <- tmbstan(m$obj, chains = 4, iter = 8000, warmup = 2000,
+mcmc_out <- tmbstan(tmb_model$obj, chains = 4, iter = 8000, warmup = 2000,
                     cores = getOption('mc.cores', 4))

### 2020-02-07 article.Rnw
Remove all comments from template
Starting to pull together implementations description.

@@
 
-%% -- Article metainformation (author, title, ...) -----------------------------
-
-%% - \author{} with primary affiliation
-%% - \Plainauthor{} without affiliations
-%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
-%% - \AND starts a new line, \And does not.
 \author{Anita K Nandi\\University of Oxford
@@
 
-%% - \title{} in title case
-%% - \Plaintitle{} without LaTeX markup (if any)
-%% - \Shorttitle{} with LaTeX markup (if any), used as running title
 \title{disaggregation: An \proglang{R} Package for Bayesian Spatial Disaggregation Modelling}
@@
 
-%% - \Abstract{} almost as usual
 \Abstract{
@@
 
-%% - \Keywords{} with LaTeX markup, at least one required
-%% - \Plainkeywords{} without LaTeX markup (if necessary)
-%% - Should be comma-separated and in sentence case.
 \Keywords{Bayesian spatial modelling, disaggregation modelling, TMB}
@@
 
-%% - \Address{} of at least one author
-%% - May contain multiple affiliations for each author
-%%   (in extra lines, separated by \emph{and}\\).
-%% - May contain multiple authors for the same affiliation
-%%   (in the same first line, separated by comma).
 \Address{
@@
 
-%% -- Introduction -------------------------------------------------------------
-
-%% - In principle "as usual".
-%% - But should typically have some discussion of both _software_ and _methods_.
-%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
-%% - If such markup is in (sub)section titles, a plain text version has to be
-%%   added as well.
-%% - All software mentioned should be properly \cite-d.
-%% - All abbreviations should be introduced.
-%% - Unless the expansions of abbreviations are proper names (like "Journal
-%%   of Statistical Software" above) they should be in sentence case (like
-%%   "generalized linear models" below).
-
 \section{Introduction} \label{sec:intro}
@@
 the number of rows in the covariate data is different to that of the response data.
-The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
+The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to be very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
 Disaggregation models can be implemented in \pkg{TMB}~\citep{tmb2016}, or template model builder, with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
 The \pkg{disaggregation} package allows this process to be streamlined, to make it easy for the user to run disaggregation models at the expense of some flexibility.
-
-%% -- Manuscript ---------------------------------------------------------------
-
-%% - In principle "as usual" again.
-%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
-%%   avoid empty lines before and after the equation (which would signal a new
-%%   paragraph.
-%% - When describing longer chunks of code that are _not_ meant for execution
-%%   (e.g., a function synopsis or list of arguments), the environment {Code}
-%%   is recommended. Alternatively, a plain {verbatim} can also be used.
-%%   (For executed code see the next section.)
 
@@
 
-\section{TMB implementation} \label{sec:implementation}
+\section{Implementation} \label{sec:implementation}
 
@@
 Evaluation of the objective function and its derivatives is performed via \proglang{R}.
+
+\subsection{Other implementations}
+
+The simplest method to disaggregate data from the polygon level to the high-resolution pixel level is to simply apply the value given by the polygon across all pixels within that polygon.
+This is the approach used by the \code{rasterize} function in the \pkg{raster} package, the \code{fasterize} function in the \pkg{fasterize} package and \code{st_interpolate_aw} from the \pkg{sf} package.
+Similarly, for time-series data there are functions such as \code{td} in the \pkg{tempdisagg} package.
+These methods are not based on a statistical model and cannot help estimate the heterogenous distribution of the disease within a polygon.
+Furthermore, these methods cannot make predictions outside of the area for which polygon data is available.
+
+The popular R package \pkg{INLA} can perform disaggregation regression in the simplest case of ...
+
+li etc.
 
@@
 <<installrun, echo=FALSE, message=FALSE>>=
-#devtools::install_github("aknandi/disaggregation")
 library(disaggregation, quietly = TRUE)
@@
 
-%% -- Bibliography -------------------------------------------------------------
-%% - References need to be provided in a .bib BibTeX database.
-%% - All references should be made with \cite, \citet, \citep, \citealp etc.
-%%   (and never hard-coded). See the FAQ for details.
-%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
-%% - Titles in the .bib should be in title case.
-%% - DOIs should be included where available.
-
 \bibliography{refs}
 
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-% \newpage
-% 
-% \begin{appendix}
-% 
-% \section{More technical details} \label{app:technical}
-% 
-% \begin{leftbar}
-% Appendices can be included after the bibliography (with a page break). Each
-% section within the appendix should have a proper section title (rather than
-% just \emph{Appendix}).
-% 
-% For more technical style details, please check out JSS's style FAQ at
-% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-% which includes the following topics:
-% \begin{itemize}
-%   \item Title vs.\ sentence case.
-%   \item Graphics formatting.
-%   \item Naming conventions.
-%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
-%   \item Trouble shooting.
-%   \item Many other potentially helpful details\dots
-% \end{itemize}
-% \end{leftbar}
-% 
-% 
-% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-% 
-% \begin{leftbar}
-% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-% formats of author-year citations and allow to include additional details (e.g.,
-% pages, chapters, \dots) in brackets. In case you are not familiar with these
-% commands see the JSS style FAQ for details.
-% 
-% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-% when acquiring the entries automatically from mixed online sources. However,
-% it is important that informations are complete and presented in a consistent
-% style to avoid confusions. JSS requires the following format.
-% \begin{itemize}
-%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-%     be used in the references.
-%   \item Titles should be in title case.
-%   \item Journal titles should not be abbreviated and in title case.
-%   \item DOIs should be included where available.
-%   \item Software should be properly cited as well. For \proglang{R} packages
-%     \code{citation("pkgname")} typically provides a good starting point.
-% \end{itemize}
-% \end{leftbar}
-% 
-% \end{appendix}
-
-%% -----------------------------------------------------------------------------
-
 \end{document}

### 2020-02-07 article.Rnw
Add refs to packages.

@@
 The simplest method to disaggregate data from the polygon level to the high-resolution pixel level is to simply apply the value given by the polygon across all pixels within that polygon.
-This is the approach used by the \code{rasterize} function in the \pkg{raster} package, the \code{fasterize} function in the \pkg{fasterize} package and \code{st_interpolate_aw} from the \pkg{sf} package.
-Similarly, for time-series data there are functions such as \code{td} in the \pkg{tempdisagg} package.
+This is the approach used by the \code{rasterize} function in the \pkg{raster} package, the \code{fasterize} function in the \pkg{fasterize} package and \code{st_interpolate_aw} from the \pkg{sf} package \citep{raster, fasterize, sf}.
+Similarly, for time-series data there are functions such as \code{td} in the \pkg{tempdisagg} package \citep{tempdisagg}.
 These methods are not based on a statistical model and cannot help estimate the heterogenous distribution of the disease within a polygon.

### 2020-02-07 article.Rnw
Add other implementations.

@@
 
-The popular R package \pkg{INLA} can perform disaggregation regression in the simplest case of ...
-
-li etc.
+A number of other implementations use statistical models to estimate high resolution disease risk from aggregated disease data.
+To the best of our knowledge, these all only include covariates at the polygon level rather than at the pixel level.
+The popular R package \pkg{INLA} can perform disaggregation regression in the simplest case of a normal likelihood, an identity link function and no high resolution covariates \citep{moraga2017geostatistical, INLA, wilson2017pointless}.
+The \pkg{SDALGCP} package fits the model using Monte Carlo Maximum Likelihood \citep{johnson2019spatially}.
+It can only fit models with a Poisson likelihood with covariates at the polygon level.
+Finally, the \pkg{lgcp} package uses a data augmentation approach to fit models to aggregated data \citep{lgcp}.  
+This Monte Carlo data augmentation approach will be very slow for large areas or large numbers of cases.
 

### 2020-02-14 article.Rnw
Update to using the new version of the package. Mainly changing plotting and function/class names

@@
 
-The main functions are \code{prepare_data}, \code{fit_model} and \code{predict}.
+The main functions are \code{prepare_data}, \code{disag_model} and \code{predict}.
 
@@
 
+\begin{figure}[!h]
+\centering
 %
-<<plotdata, eval=FALSE>>=
+<<plotdata, fig.width = 15, fig.height = 15>>=
 plot(dis_data)
@@
 %
+\caption{\label{fig:plotdata} Maps of Madagascar showing the data used in the disaggregation model.
+These plots are produced when calling the \code{plot} function on the \code{disag\_data} object.
+The plots show the number of malaria cases in each administrative unit (top left), maps of the four covariates used in the model (top right), and inla.mesh object that will be used to make the spatial field (bottom).}
+\end{figure}
 
-%
-<<plotdatahidden, echo=FALSE, results='hide', fig.show='hide'>>=
-data_plots <- plot(dis_data)
-@
-%
+The \code{prepare_data} function returns an object of class \code{disag_data}, which is designed to be used directly in the \code{disag_model} function.
 
-%
-<<subplotdata, fig.cap='Maps of Madagascar showing the data used in the disaggregation model.
-These plots are produced when calling the plot function on the disag.data object.
-The plots show: (a) number of malaria cases in each administrative unit, (b) maps of the four covariates used in the model; and (c) inla.mesh object that will be used to make the spatial field.', fig.subcap=c('', '', ''), out.width='0.5\\linewidth', echo=FALSE, fig.ncol = 2, dev = 'png', fig.width = 4, fig.height = 5, fig.pos = 'H'>>=
-data_plots$polygon
-data_plots$covariates
-INLA::plot.inla.mesh(dis_data$mesh, main = 'INLA mesh for spatial field')
-@
-%
-
-The \code{prepare_data} function returns an object of class \code{disag.data}, which is designed to be used directly in the \code{fit_model} function.
-
-Now we can fit the disaggregation model using \code{fit_model}.
+Now we can fit the disaggregation model using \code{disag_model}.
 Here we use a Poisson likelihood for the incident count data with a log link function.
@@
 %
-<<fit_model, message = FALSE>>=
-fitted_model <- fit_model(data = dis_data,
+<<disag_model, message = FALSE>>=
+fitted_model <- disag_model(data = dis_data,
                             iterations = 1000,
@@
 \caption{\label{fig:fitmodel} Plot summarising the results of the fitted model.
-These plots are produced when calling the \code{plot} function on the \code{fit.result} object.
+These plots are produced when calling the \code{plot} function on the \code{disag\_model} object.
 The fixed effects plot (left) shows the fitted parameter values with uncertainty estimation for all the fixed effects in the model.
@@
 
-The \code{fit_model} function returns an object of class \code{fit.result}, which is designed to be used directly in the \code{predict} function.
+The \code{disag_model} function returns an object of class \code{disag_model}, which is designed to be used directly in the \code{predict} function.
 Therefore, now that we have fitted the model, we are ready to predict the malaria incidence rate across Madagascar.
@@
 <<predict_model>>=
-model_predictions <- predict(fitted_model)
+model_prediction <- predict(fitted_model)
 @
@@
 
-The function \code{predict} returns a list of two objects: the mean predictions (of class \code{predictions}) and the uncertainty rasters (of class \code{uncertainty}).
+The function \code{predict} returns a object of class \code{disag_prediction} containing a list of two objects: the mean predictions and the uncertainty rasters.
 The mean predictions contain a raster of the mean prediction of the incidence rate, as well as rasters of the field, iid (if predicted) and covariate component of the linear predictor.
-The \code{uncertainty} object contains a RasterStack of the prediction realisations and a RasterStack of the upper and lower credible intervals.
+The uncertainty predictions contains a RasterStack of the prediction realisations and a RasterStack of the upper and lower credible intervals.
 
-The \code{plot} function can be used on both the \code{predictions} and \code{uncertainty} objects, as shown in Figures~\ref{fig:subpredictmean} and \ref{fig:predictuncertainty} respectively.
-
-%
-<<plotpredictions, eval=FALSE>>=
-plot(model_predictions$mean_predictions)
-@
-%
-
-%
-<<hiddenplotmean, echo=FALSE, results='hide', fig.show='hide'>>=
-mean_predictions <- plot(model_predictions$mean_predictions)
-@
-%
-
-%
-<<subpredictmean, fig.cap='Maps of Madagascar showing the fine-scale predictions from the disaggregation model.
-These maps are produced when calling the plot function on the predictions object.
-The maps show the (a) mean prediction of malaria incidence rate, (b) covariate contribution to the linear predictor, and (c) field contribution to the linear predictor.', fig.subcap=c('', '', ''), out.width='0.33\\linewidth', echo=FALSE, dev = 'png', fig.width = 4, fig.height = 6, fig.pos = 'H'>>=
-mean_predictions$mean
-mean_predictions$covariates
-mean_predictions$field
-@
-%
+The \code{plot} function can be used on the \code{disag_prediction} objects, as shown in Figure \ref{fig:plotprediction}.
 
@@
 %
-<<plotuncertainty, dev = 'png', fig.width = 10, fig.height = 5>>=
-plot(model_predictions$uncertainty_predictions)
+<<plotprediction, dev = 'png', fig.width = 15, fig.height = 10>>=
+plot(model_prediction)
 @
 %
-\caption{\label{fig:predictuncertainty} Maps of Madagascar showing the fine-scale predictions of lower (2.5\%) and upper (97.5\%) credible intervals of the malaria incidence rate from the disaggregation model.
-These maps are produced when calling the \code{plot} function on the \code{uncertainty} object.}
+\caption{\label{fig:plotprediction} Maps of Madagascar showing the fine-scale predictions of mean, lower (2.5\%) and upper (97.5\%) credible intervals of the malaria incidence rate from the disaggregation model.
+These maps are produced when calling the \code{plot} function on the \code{disag\_prediction} object.}
 \end{figure}
@@
 Here we fit the same model by running MCMC using the \pkg{tmbstan} package, with the NUTS (no-u-turn sampler) algorithm \citep{hoffman2014no} using four chains.
-It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{fit_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
+It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{disag_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
 The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 94 hours.
@@
 
-tmb_model <- fit_model(data = dis_data, 
+tmb_model <- disag_model(data = dis_data, 
                          iterations = 1,

### 2020-02-14 article.Rnw
Commas after i.e. and e.g.

@@
 
-Disaggregation modelling is unorthodox as the predictions are at a different scale to the response data, i.e.
-the number of rows in the covariate data is different to that of the response data.
+Disaggregation modelling is unorthodox as the predictions are at a different scale to the response data, i.e., the number of rows in the covariate data is different to that of the response data.
 The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to be very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
@@
 The process that is being counted occurs in continuous space that we model as a high-resolution, square lattice for convenience.
-The data, $y_i$,  are assumed to be created by the aggregation of the counts over the polygon, i.e.
-the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
+The data, $y_i$,  are assumed to be created by the aggregation of the counts over the polygon, i.e., the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
 The rate is defined such that the number of cases in this pixel can be calculated by multiplying the rate by the aggregation raster.
@@
 In this case our pixel predictions are malaria incidence rate, so we use the population raster to aggregate pixel incidence rate by summing the number of cases (rate weighted by population).
-If no aggregation raster is provided a uniform distribution is assumed, i.e.
-the pixel predictions are aggregated to polygon level by summing the pixel values unaltered.
+If no aggregation raster is provided a uniform distribution is assumed, i.e., the pixel predictions are aggregated to polygon level by summing the pixel values unaltered.
 
@@
 
-In order to know which pixels (i.e.
-which rows) are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
+In order to know which pixels (i.e., which rows) are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
 Additionally, an INLA mesh is built to use for the spatial field (\code{build_mesh} function).
@@
 For the uncertainty calculations, parameter values are sampled from the posterior distribution and summarised.
-The number of parameter draws used to calculate the uncertainty is set by the user via the \code{N} parameter (default: 100), and the size of the credible interval (e.g.
-75\%, 95\%) to be calculated when summarising is set via the argument \code{CI} (default: 0.95).
+The number of parameter draws used to calculate the uncertainty is set by the user via the \code{N} parameter (default: 100), and the size of the credible interval (e.g., 75\%, 95\%) to be calculated when summarising is set via the argument \code{CI} (default: 0.95).
 

### 2020-02-14 article.Rnw
Removed all non-jss style formatting

@@
 Disaggregation regression models use response data aggregated over large heterogenous regions to make predictions at fine-scale over the region by using fine-scale covariates to inform the heterogeneity.
-This paper presents the \proglang{R} package \emph{disaggregation}, which provides functionality to streamline the process of running a disaggregation model for fine-scale predictions.
+This paper presents the \proglang{R} package \pkg{disaggregation}, which provides functionality to streamline the process of running a disaggregation model for fine-scale predictions.
 }
@@
 In this section we show how to use the \pkg{disaggregation} package using a dataset of aggregated malaria case counts across Madagascar.
-Malaria is an infectious disease caused by parasites of the \emph{Plasmodium} group, transmitted by \emph{Anopheles} mosquitoes.
+Malaria is an infectious disease caused by parasites of the Plasmodium group, transmitted by Anopheles mosquitoes.
 Malaria transmission is therefore closely related to mosquito and parasite development.

### 2020-02-14 article.Rnw
Add \pkg, \proglang, \code formatting for titles and references

@@
 
-\title{disaggregation: An \proglang{R} Package for Bayesian Spatial Disaggregation Modelling}
+\title{\pkg{disaggregation}: An \proglang{R} Package for Bayesian Spatial Disaggregation Modelling}
 \Plaintitle{disaggregation: An R Package for Bayesian Spatial Disaggregation Modelling}
@@
 \Keywords{Bayesian spatial modelling, disaggregation modelling, TMB}
-\Plainkeywords{Bayesian spatial modelling, disaggregation modelling, TMB}
+\Plainkeywords{Bayesian spatial modelling, disaggregation modelling, \pkg{TMB}}
 

### 2020-02-18 article.Rnw

@@
   Oxford, UK\\
-  E-mail: \email{anita.nandi@bdi.ox.ac.uk}
+  E-mail: \email{anita.k.nandi@gmail.com}
 }

### 2020-02-24 article.Rnw
Remove setting of priors, fix figure refs, remove LST sd

@@
 Therefore, such environmental covariates would be useful in a malaria disaggregation model to inform fine-scale distributions.
-In this model we use the environmental covariates of mean land surface temperature (LST), variation in LST, elevation, and enhanced vegetation index (EVI) to inform spatial heterogeneity in malaria risk.
+In this model we use the environmental covariates of mean land surface temperature (LST), elevation, and enhanced vegetation index (EVI) to inform spatial heterogeneity in malaria risk.
 These covariates are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
@@
 The summary function returns information on how many pixels and polygons the data contains, how many pixels in the smallest and largest polygons and a summary of the covariate data.
-The \code{plot} functions plots a map of the polygon response data, the covariate rasters and the INLA mesh, as shown in Figure~\ref{fig:subplotdata}.
+The \code{plot} functions plots a map of the polygon response data, the covariate rasters and the INLA mesh, as shown in Figure~\ref{fig:plotdata}.
 
@@
 The spatial field and iid effect are components of the model by default, they can be turned off using the \code{field} and \code{iid} flags.
-We specify all of the priors for the model in a single list.
-Hyperpriors for the field are given as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{min}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{min}) = \sigma_{prob}$.
-Similarly, the user specifies penalised complexity priors for the iid effect.
+In this example we use the default priors.
+Hyperpriors for the field are implemented as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{min}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{min}) = \sigma_{prob}$.
+Similarly, penalised complexity priors are implemented for the iid effect.
 The \code{iterations} argument specifies the maximum number of iterations the model can run for to find an optimal point.
@@
                             family = 'poisson',
-                            link = 'log',
-                            priors = list(priormean_intercept = 0,
-                                          priorsd_intercept = 2,
-                                          priormean_slope = 0.0,
-                                          priorsd_slope = 0.4,
-                                          prior_rho_min = 3,
-                                          prior_rho_prob = 0.01,
-                                          prior_sigma_max = 1,
-                                          prior_sigma_prob = 0.01))
+                            link = 'log')
 @
@@
 The \code{summary} function gives the estimate and standard error of the fixed effect parameters in the model, as well as the negative log likelihood and in-sample performance metrics (root mean squared error, mean absolute error, pearson correlation coefficient and spearman rank correlation coefficient).
-The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions, as shown in Figure~\ref{fig:fitmodel}.
+The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions, as shown in Figure~\ref{fig:plotmodel}.
 
@@
 %
-\caption{\label{fig:fitmodel} Plot summarising the results of the fitted model.
+\caption{\label{fig:plotmodel} Plot summarising the results of the fitted model.
 These plots are produced when calling the \code{plot} function on the \code{disag\_model} object.
@@
 
-tmb_model <- disag_model(data = dis_data, 
-                         iterations = 1,
+model_object <- make_model_object(data = dis_data,
                                   family = 'poisson',
-                         link = 'log',
-                         priors = list(priormean_intercept = 0,
-                                       priorsd_intercept = 2,
-                                       priormean_slope = 0.0,
-                                       priorsd_slope = 0.4,
-                                       prior_rho_min = 3,
-                                       prior_rho_prob = 0.01,
-                                       prior_sigma_max = 1,
-                                       prior_sigma_prob = 0.01))
+                                  link = 'log')
 
-# Running the MCMC algorithm for 94 hours
+# Running the MCMC algorithm - takes a long time
 start <- Sys.time()
-mcmc_out <- tmbstan(tmb_model$obj, chains = 4, iter = 8000, warmup = 2000,
+mcmc_out <- tmbstan(model_object, chains = 4, iter = 8000, warmup = 2000,
                     cores = getOption('mc.cores', 4))
@@
 
-# Plot the trace of the parameter values for each sampling method
+# Plot the trace of the parameter values
 stan_trace(mcmc_out, 

### 2020-02-24 article.Rnw

@@
 <<kintroptions, echo=FALSE>>=
+knitr::knit_hooks$set(render_sweave())
 knitr::opts_chunk$set(
   collapse = TRUE,
-  comment = "#>",
   cache = TRUE,
@@
   fig.height = 4,
-  fig.pos = 'H'
+  fig.pos = 'H',
+  prompt=TRUE
 )
+options(replace.assign=TRUE, width=90, prompt="R> ")
 @

### 2020-02-25 reproduction_script.R
Neaten reproduction script

@@ -1,9 +1,20 @@
+# disaggregation: An R Package for Bayesian Spatial Disaggregation 
+# Modelling
+#
+# This replication script runs all code needed to reproduce results
+# in the paper above
+#
+# If you want to re-run the MCMC time comparison for full replicability, 
+# please set the variable (run_mcmc) below to TRUE.
+# NOTE: running the MCMC will take a long time even on high
+# performance systems.
+
+run_mcmc <- FALSE
 
 #######
 # Example predicting malaria in Madagascar
 #######
 
-devtools::install.packages("disaggregation")
 library(disaggregation)
 library(raster)
 library(tmbstan)
@@ -14,7 +25,7 @@ covariate_stack <- getCovariateRasters('data/covariates',
                                        shape = population_raster)
 
 dis_data <- prepare_data(polygon_shapefile = shapes, 
-                         covariate_rasters = covariate_stack[[1:3]], 
+                         covariate_rasters = covariate_stack, 
                          aggregation_raster = population_raster, 
                          mesh.args = list(max.edge = c(0.7, 8), 
                                           cut = 0.05, 
@@ -46,11 +57,6 @@ plot(model_prediction)
 ##########
 # Comparison with MCMC
 ##########
-
-# The MCMC takes a very long time to converge (96 hours)
-# If you want to run this set the run_mcmc flag to true
-run_mcmc <- FALSE
-
 
 if(run_mcmc) {
   

### 2020-02-25 article.Rnw
Minor paper edits

@@
 becoming increasingly sought after in disease risk mapping~\citep{li2012log, diggle2013spatial, wilson2017pointless}, especially in malaria mapping~\citep{sturrock2014fine, sturrock2016mapping}. 
-Disaggregation regression, first applied in species distribution modelling in ecology~\citep{keil2013downscaling, barwell2014can} has now become an important method in disease risk mapping \citep{falciparum2019, vivax2019}.
+Disaggregation regression, first applied in species distribution modelling in ecology~\citep{keil2013downscaling} has now become an important method in disease risk mapping \citep{falciparum2019, vivax2019}.
 The aggregation of response data over large heterogenous regions is problematic for making fine-scale predictions, as relationships learned between variables at one spatial scale may not hold at other scales \citep{wakefield2006health}.
@@
 \end{equation}
-which shrinks towards a base model of no polygon-specific effect. The $\sigma_{u, max}$ and $\sigma_{u, prob}$ are set by the user.
+which shrinks towards a base model of no polygon-specific effect. The values $\sigma_{u, max}$ and $\sigma_{u, prob}$ are set by the user.
 
@@
 Malaria transmission is therefore closely related to mosquito and parasite development.
-Environmental factors such as temperature, rainfall and elevation have been shown to have signifcant effect on mosquito survival and development; in general mosquitoes favour warm, humid environments with moderate rainfall.
+Environmental factors such as temperature, rainfall and elevation have been shown to have signifcant effects on mosquito survival and development; in general mosquitoes favour warm, humid environments with moderate rainfall.
 Therefore, such environmental covariates would be useful in a malaria disaggregation model to inform fine-scale distributions.

### 2020-02-25 article.Rnw
Tiny reword

@@
 )
-options(replace.assign=TRUE, width=90, prompt="R> ")
+options(prompt="R> ")
 @
@@
 The \pkg{disaggregation} package contains a C++ function that defines the model and computes the joint likelihood as a function of the parameters and the random effects, in the format expected by \pkg{TMB}.
-The \pkg{TMB} package then calculates estimates of both parameters and random effects using the Laplace approximation for the likelihood.
+The \pkg{TMB} package then calculates estimates of both parameters and random effects using a Laplace approximation.
 Evaluation of the objective function and its derivatives is performed via \proglang{R}.

### 2020-02-25 article.Rnw
Tiny edits

@@
 
-# Running the MCMC algorithm - takes a long time
 start <- Sys.time()
@@
 
-# Plot the trace of the parameter values
 stan_trace(mcmc_out, 
            pars = c('intercept', 
-                    'slope[1]', 'slope[2]', 'slope[3]', 'slope[4]', 
+                    'slope[1]', 'slope[2]', 'slope[3]',
                     'iideffect_log_tau', 'log_sigma', 'log_rho'))
@@
 install.packages("INLA", 
-                 repos = c(getOption("repos"), 
+                 repos=c(getOption("repos"), 
                          INLA="https://inla.r-inla-download.org/R/stable"), 

### 2020-02-27 article.Rnw
Minor text edits

@@
 where $\beta$ are regression coefficients, $X_{ij}$ are covariate values, GP is a Gaussian random field and $u_i$ is a polygon-specific iid effect.
-The user-defined link function is typically log, identity and logit for Poisson, Normal and Binomial likelihoods, respectively.
+The user-defined link function is typically log, identity and logit for Poisson, Normal and binomial likelihoods, respectively.
 The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
@@
 \begin{equation} \label{eq:poisson}
-y_i \sim \textrm{Pois}(\textrm{cases}_i)
+y_i \sim \textrm{Poisson}(\textrm{cases}_i)
 \end{equation}
@@
 In the example of disease mapping, Poisson or Gaussian likelihoods could be used when the quantity observed, $y_i$, is the total number of cases in a given polygon.
-The Binomial model could be used when $y_i$ is the prevalence of a disease in a sample of $M_i$ people in the polygon.
+The binomial model could be used when $y_i$ is the prevalence of a disease in a sample of $M_i$ people in the polygon.
 
@@
 These covariates are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
-Malaria incidence rate is often given per thousand cases and given the term annual parasite index (API).
+Malaria incidence rate is often given per thousand people per year given the term annual parasite index (API).
 
@@
 The aggregation raster defines how the pixels within each polygon are aggregated.
-The disaggregation model performs a weighted sum of the pixel prediction, weighted by the pixel values in the aggregation raster.
+The disaggregation model performs a weighted sum of the pixel predictions, weighted by the pixel values in the aggregation raster.
 In this case our pixel predictions are malaria incidence rate, so we use the population raster to aggregate pixel incidence rate by summing the number of cases (rate weighted by population).
@@
 Ideally the NAs in the data would be dealt with by the user beforehand, however, setting na.action = TRUE will automatically deal with NAs.
-It removes any polygons that have NAs as a response, sets any aggregation pixels with NA to zero and sets covariate NAs pixels to the median value for that covariate across all polygons.
+It removes any polygons that have NAs as a response, sets any aggregation pixels with NA to zero and sets covariate NA pixels to the median value for that covariate across all polygons.
 
@@
 Now we can fit the disaggregation model using \code{disag_model}.
-Here we use a Poisson likelihood for the incident count data with a log link function.
+Here we use a Poisson likelihood for the incidence count data with a log link function.
 Options for the likelihood are Gaussian, Poisson and binomial, and options for the link function are logit, log and identity.

### 2020-02-27 article.Rnw
typo on sigma_max

@@
 In this example we use the default priors.
-Hyperpriors for the field are implemented as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{min}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{min}) = \sigma_{prob}$.
+Hyperpriors for the field are implemented as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{max}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
 Similarly, penalised complexity priors are implemented for the iid effect.

### 2020-03-02 article.Rnw
Updated paper with new MCMC (3 covariates)

@@
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{disag_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
-The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 94 hours.
+The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 71 hours.
 This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations, doubling the number of iterations each time until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
@@
 \begin{tabular}{l|ll|ll}
- & \multicolumn{2}{|c|}{MCMC (94 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
+ & \multicolumn{2}{|c|}{MCMC (71 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
 Parameter & Mean & SD & Mean & SD \\
 \hline
-Intercept & -3.09 & 0.33 & -3.12 & 0.27 \\
-Slope 1 & 0.20 & 0.28 & 0.24 & 0.27 \\
-Slope 2 & 0.32 & 0.21 & 0.30 & 0.21\\
-Slope 3 & -0.36 & 0.20 & -0.37 & 0.20 \\
-Slope 4 & -0.24 & 0.23 & -0.24 & 0.21 \\
-$log(\tau_u)$ & 1.08 & 0.27 & 1.10 & 0.27 \\
-$log(\sigma)$ & -3.34 & 0.59 & -3.55 & 0.59 \\
-$log(\rho)$ & 0.72 & 0.31 & 0.56 & 0.30 \\
+Intercept & -3.22 & 0.33 & -3.20 & 0.27 \\
+Slope 1 & -0.47 & 0.20 & -0.47 & 0.20 \\
+Slope 2 & 0.41 & 0.22 & 0.39 & 0.22 \\
+Slope 3 & 0.21 & 0.29 & 0.21 & 0.29 \\
+$log(\tau_u)$ & 0.90 & 0.27 & 0.98 & 0.27 \\
+$log(\sigma)$ & 0.062 & 0.009 & -0.002 & 0.166 \\
+$log(\rho)$ & 0.82 & 0.35 & 0.63 & 0.33 \\
 \end{tabular}
@@
 The \pkg{disaggregation} package produces similar results to the MCMC algorithm.
-However the model fitting using the \pkg{disaggregation} package took 56 seconds in contrast to the 94 hours taken for the MCMC run.
+However the model fitting using the \pkg{disaggregation} package took 56 seconds in contrast to the 71 hours taken for the MCMC run.
 Therefore, it can be seen that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that can be prohibitively slow using MCMC.
@@
 \centering
-\includegraphics[width=\linewidth]{figures/mcmc_long_trace.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 94 hours. It was run for 8000 iterations with 2000 of those as warmup.}
+\includegraphics[width=\linewidth]{figures/mcmc_trace.pdf}
+\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 71 hours. It was run for 8000 iterations with 2000 of those as warmup.}
 \label{fig:mcmclong}

### 2021-01-03 article.Rnw
Updates from review comments

@@
 <<kintroptions, echo=FALSE>>=
-knitr::knit_hooks$set(render_sweave())
 knitr::opts_chunk$set(
@@
 \begin{document}
+%\SweaveOpts{concordance=TRUE}
 
@@
 The data, $y_i$, are assumed to be created by the aggregation of the counts over the polygon, i.e., the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
-The rate is defined such that the number of cases in this pixel can be calculated by multiplying the rate by the aggregation raster.
+As well as the model covariates, an aggregation raster can be provided transform from the pixel level predictions (rate) to count data, from which it is trivial to calculate the corresponding polygon value by summing up the values for the individual pixels within the polygon.
 For example, in epidemiology, we may have as our response the number of people that contract a certain disease in a given period of time (case incidence).
 Our rate would be the incidence rate of cases per population, where the aggregation raster corresponds to population.
+Predictions of incident rate per population are made at pixel level, the population (aggregation) raster is used to convert these values to number of cases, which can then be summed to get predicted polygon level cases.
 
@@
 The user-defined link function is typically log, identity and logit for Poisson, Normal and binomial likelihoods, respectively.
-The Gaussian random field has a Matern covariance function parameterised by $\rho$, the range (approximately the distance beyond which correlation is less than 0.1), and $\sigma$, the marginal standard deviation.
+The Gaussian random field has a Matern covariance function, defined by:
+
+%
+\begin{equation}
+Cov(d) = \frac{\sigma^2}{\Gamma(\nu)2^{\nu - 1}}(\kappa d)^{\nu}K_{\nu}(\kappa d)
+\end{equation}
+%
+
+with two hyperparameters: $\rho = \frac{\sqrt{8\nu}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating to the magnitude of the variation in the field.
+The parameter $\nu$ is the smoothness and is fixed at 2.
+The Matern field is at least $m$ times mean square differentiable if and only if $\nu > m$~\citep{stein1999interpolation}, therefore $\nu = 2$ is chosen as it is the smallest integer (least smooth) for the field to be differentiable.
+The Gamma function is given by $\Gamma(\nu) = (\nu - 1)!$ and $K_{\nu}$ is the modified Bessel function of the second kind.
+
 As we are working in a Bayesian setting, each of the model parameters and hyperparameters are given a prior, which is discussed later.
 
-The pixel predictions are then aggregated to the polygon level using the weighted sum (via the aggregation raster $a_{ij}$):
+The predictions are made at pixel level, however the response data exists as count data at the polygon level.
+Therefore, to calculate the contribution  to the likelihood from the data we must aggregate the pixel predictions, (via the aggregation raster $a_{ij}$, to get the predicted polygon count.
 
@@
 %
+where $N_i$ is the total number of pixels in polygon $i$.
 The different likelihoods correspond to slightly different models ($y_i$ is the response count data):
@@
 The regression parameters and intercept are given Gaussian priors, where the default priors are $\beta_0 \sim N(0, 2)$ and $\beta_i \sim N(0, 0.4)$.
-A penalised complexity prior is placed on the scale and range parameters of the random field \citep{fuglstad2018constructing} such that
+It is expected that some of the spatial variation can be described by the covariates, and the Gaussian field can help describe the remaining spatial variation that is missing from the covariate information.
+Therefore, the priors on the covariates can be set to allow the covariates to explain some, but not all, of the variation in the response data.
+For the Gaussian random field, penalised complexity priors are used, which are constructed to  penalise against deviating from the simpler base model, which in this case is a flat field, i.e.
+zero variance and infinite range~\citep{fuglstad2018constructing}.
+A penalised complexity prior is placed on the scale and range parameters of the random field such that
 \begin{align}
@@
 where the values $\rho_{min}, \,\rho_{prob}, \,\sigma_{max}, \,\sigma_{prob}$ are set by the user.
+The default values for these parameters within the package are driven by the nature of the data provided in the model.
+The default prior for $\rho_{min}$ is set at a third of the spatial area covered by the polygons, and $\rho_{prob}$ is set at 0.1.
+The default prior for $\sigma_{max}$ is set to the standard deviation of the normalised response data, and $\sigma_{prob}$ is set at 0.1.
+
+The joint penalised complexity prior, $\pi$, corresponding to a base model with infinite range and zero variance~\citep{fuglstad2018constructing} is given by:
+
+%
+\begin{equation} 
+log\left(\pi(\sigma, \rho)\right) = log\left(\tilde{\lambda_1}\right) +  log\left(\tilde{\lambda_2}\right) - 2log(\rho) - \frac{\tilde{\lambda_1}}{\rho} - \tilde{\lambda_2} \sigma
+\end{equation}
+
+where
+
+\begin{equation} 
+\tilde{\lambda_1} = -\rho_{min}log(\rho_{prob}) \qquad \text{and} \qquad \tilde{\lambda_2} = -\frac{log(\sigma_{prob})}{\sigma_{max}} 
+\end{equation}
+%
+
 This prior shrinks the field towards a base model with zero variance and infinite range, in other words regularising towards a flatter field with smaller magnitude.
+
 The polygon-specific effects $u_1,...,u_N$ have Gaussian priors centered at 0 with standard deviation $\sigma_{u}$ (where the precision $\tau_{u} = 1/\sigma_{u}^2$).
-A penalised complexity prior \citep{simpson2017penalising} is placed on $\sigma_{u}$ such that
+A penalised complexity prior is placed on $\sigma_{u}$~\citep{simpson2017penalising} such that
 \begin{equation}
@@
 \end{equation}
-which shrinks towards a base model of no polygon-specific effect. The values $\sigma_{u, max}$ and $\sigma_{u, prob}$ are set by the user.
+where values $\sigma_{u, max}$ and $\sigma_{u, prob}$ are set by the user.
+The penalised complexity prior, $\pi_u$, corresponding to a base model with no polygon-specific effect~\citep{simpson2017penalising} is given by:
 
-For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau \sim \log\Gamma(\text{shape} = 1, \text{rate} = 5 \times 10^{-5})$, to regularise the dispersion ($\sigma = 1/\sqrt\tau$) to take low values.
-This is the same as the prior set by INLA for the dispersion of the normal likelihood.
+%
+\begin{equation} 
+log\left(\pi(\lambda)\right) = log\left(\frac{\lambda}{2}\right) - \frac{3}{2}log(\tau_u) - \frac{\lambda}{\sqrt{\tau_u}} 
+\end{equation}
+
+where
+
+\begin{equation} 
+\lambda = -\frac{log(\sigma_{u, prob})}{\sigma_{u, max}} \qquad \text{and} \qquad \tau_u = \frac{1}{\sigma_u^2}
+\end{equation}
+%
+
+This prior shrinks towards a base model of no polygon-specific effect.
+
+For models that use a Gaussian likelihood, a log gamma prior is set on the log of the precision, $\log\tau_u \sim \log\Gamma(\text{shape} = 1, \text{rate} = 5 \times 10^{-5})$, to regularise the dispersion, $\sigma_u$, to take low values.
+This is chosen to be consistent with the prior set by INLA for the dispersion of the normal likelihood.
 
@@
 Evaluation of the objective function and its derivatives is performed via \proglang{R}.
+
+\pkg{TMB} can fit both Frequentist and Bayesian models. In this case we have chosen the Bayesian approach as it is most widely used in the spatial modelling community.
 
@@
 
-In this section we show how to use the \pkg{disaggregation} package using a dataset of aggregated malaria case counts across Madagascar.
+In this section we show how to use the \pkg{disaggregation} package using a dataset of aggregated malaria case counts across Madagascar in 2016.
 Malaria is an infectious disease caused by parasites of the Plasmodium group, transmitted by Anopheles mosquitoes.
@@
 These covariates are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
+
 Malaria incidence rate is often given per thousand people per year given the term annual parasite index (API).
+In this case our pixel predictions correspond to malaria incidence rate, so we use population to aggregate pixel incidence rate by summing the number of cases (rate weighted by population).
+Raster surfaces of population for the years 2010 and 2015, were created using data from WorldPop~\citep{tatem2017worldpop} and from GPWv4~\citep{GpwNasa} where WorldPop did not have values.
+The population raster for 2016 was created by linear interpolation between 2010 and 2015 and extending out to 2016.
+This interpolation method results in a small amount of uncertainty in population raster, however this is expected to be negligible compared to the model uncertainty.
+
+Covariate rasters or population rasters may be accompanied by significant uncertainty.
+There is no established method within the \pkg{disaggregation} package to deal with this.
+In order to propagate this uncertainty to the predictions, users may run the model many times, varying these rasters within the known uncertainty.
+Both the range in predictions for different models and the model uncertainty within a model allows a quantification of the overall uncertainty of predictions.
+
+The covariate rasters and aggregation raster provided must be of the same spatial scale and must be spatially aligned.
+For this example the rasters used are at a resolution of approximately $5\times5~km$ across Madagascar.
+For spatially misaligned rasters, there are packages in \prolang{R} to align rasters such as \code{align_rasters} in the \pkg{gdalUtils} package~\cite{gdalutils}.
 
@@
 %
-<<installrun, echo=FALSE, message=FALSE>>=
+<<installrun, echo=FALSE, message=FALSE, warnings=FALSE>>=
 library(disaggregation, quietly = TRUE)
@@
 We then read in the data to use in the \pkg{disaggregation} package.
+The \code{shapefile} function and \code{raster} function are contained within the \pkg{raster} package, they are functions to read spatial data and raster data respectively.
+These functions return objects of type SpatialPolygonsDataFrame and RasterLayer respectively.
+A SpatialPolygonsDataFrame is an \prolang{R} class that holds spatial data in the form of polygons with attributes, in this case the attributes are number of malaria cases within the polygon and polygon ID.
+A RasterLayer is an \prolang{R} class that holds a single raster.
+The function \code{getCovariateRasters} is a helper function contained within the \pkg{disaggregation} package.
+The user provides a directory name and shapefile, it finds all the raster files within that directory and creates a RasterStack with the same extend as the shapefile provided.
+In this example we are building a RasterStack of the covariate rasters with the same extent as the population raster.
+A RasterStack is a collection of RasterLayer objects with the same spatial extent and resolution.
 
@@
 This function performs various data manipulation tasks to create objects that are necessary for fitting the model.
-The required input data for the \code{prepare_data} function are a SpatialPolygonsDataFrame containing the response data and a RasterStack of covariates to be used in the model.
-The variable names in the SpatialPolygonsDataFrame of the response count data and the polygon ID are defined by the user.
+The required input data for the \code{prepare_data} function are:
+\begin{itemize}
+\item \code{polygon_shapefile}: SpatialPolygonsDataFrame containing the response data. It must contain IDs and response data.
+\item \code{covariate_raster}: RasterStack of covariates to be used in the model.
+\item \code{aggregation_raster}: RasterLayer used as the weights to aggregate the pixel values within a polygon.
+\item \code{mesh.args}: List of parameters to control the mesh used for the Gaussian Field component.
+\item \code{id_var}: the variable name of the ID variable in the \code{polygon_shapefile}.
+\item \code{response_var}: the variable name of the response variable in the \code{polygon_shapefile}.
+\item \code{na.action}: Boolean. Whether to deal with NAs or not.
+\item \code{ncores}: Number of cores to perform the parallel extraction over.
+\end{itemize}
 
@@
 The values extracted from the aggregation raster are returned as an array of values, one for each pixel.
+In order to know which pixels (i.e., which rows) are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
 
-In order to know which pixels (i.e., which rows) are contained in each polygon, a matrix is constructed that contains the start and end pixel index for each polygon (\code{getStartendindex} function).
-Additionally, an INLA mesh is built to use for the spatial field (\code{build_mesh} function).
-The \code{mesh.args} argument allows the user to supply a list of INLA mesh parameters to control the mesh used for the spatial field.
-The mesh can take several minutes to construct, so to prepare the data without buidling the mesh the user can set the \code{makeMesh} flag to FALSE.
+To fit a model with a Gaussian field we approximate a Gaussian field using Gaussian Random Markov Fields (GRMF), and solve using the stochastic partial differential equation (SPDE) approach~\citep{Lindgren2011}.
+This reduces the complexity of the problem to a sparse precision matrix, allowing explicit computation of the GMRF representation.
+This approach requires building a mesh, i.e.
+splitting the space into finite elements, and calculating field values at the mesh nodes.
+If a mesh is too fine, the field values take too long to compute, however if the mesh if too coarse, the resulting field is a poor approximation and leads to mesh artifacts in the predictions.
+The 2D mesh for the spatial field is built using the \code{build_mesh} function, which makes use of the \pkg{INLA} function \code{inla.mesh.2d}.
+The user can control the parameters of the mesh, including the granularity, using the argument \code{mesh.args} in the \code{build_mesh} function.
+The parameters \code{max.edge}, \code{cut} and \code{offset} can be set in the \code{mesh.args} list and are defined within the \code{inla.mesh.2d} function.
+By providing a two values to the \code{max.edge} parameter, the mesh contains an inner region of finer mesh, and an outer coarse region.
+This approach allows the region of interest to have a fine a mesh as necessary without also requiring time consuming computations in the sea regions.
+The mesh can take several minutes to construct, so to prepare the data without building the mesh the user can set the \code{makeMesh} flag to FALSE.
 However, it would then not be possible to fit the disaggregation model without the mesh.
@@
 
-Now we can fit the disaggregation model using \code{disag_model}.
+Now we can fit the disaggregation model using \code{disag_model}. The required inputs for the \code{disag_model} function are:
+\begin{itemize}
+\item \code{data}: Object of class \code{disag_data} returned by the \code{prepare_data} function.
+\item \code{priors}: List of priors to use for the model. For any not set, the default priors will be used.
+\item \code{iterations}: Maximum number of iterations the model can run for to find an optimal point.
+\item \code{family}: Likelihood function. Options are gaussian, poisson and binomial.
+\item \code{link}: Link function used in the model. Options are logit, log and identity. This would typically be log, identity and logit for Poisson, Normal and binomial likelihoods, respectively.
+\end{itemize}
+
 Here we use a Poisson likelihood for the incidence count data with a log link function.
-Options for the likelihood are Gaussian, Poisson and binomial, and options for the link function are logit, log and identity.
 The spatial field and iid effect are components of the model by default, they can be turned off using the \code{field} and \code{iid} flags.
 In this example we use the default priors.
-Hyperpriors for the field are implemented as penalised complexity priors - specify $\rho_{min}$ and $\rho_{prob}$ for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{max}$ and $\sigma_{prob}$ for the variation of the field, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
-Similarly, penalised complexity priors are implemented for the iid effect.
-The \code{iterations} argument specifies the maximum number of iterations the model can run for to find an optimal point.
+Hyperpriors for the field are implemented as penalised complexity priors - specify $\rho_{min}$ (\code{prior_rho_min}) and $\rho_{prob}$ (\code{prior_rho_prob}) for the range of the field, where $P(\rho < \rho_{min}) = \rho_{prob}$, and $\sigma_{max}$ (\code{prior_sigma_max}) and $\sigma_{prob}$ (\code{prior_sigma_prob}) for the variation of the field, where $P(\sigma > \sigma_{max}) = \sigma_{prob}$.
+Similarly, penalised complexity priors are implemented for the iid effect (\code{prior_iideffect_sd_max} and \code{prior_iideffect_sd_prob}).
+Any of the priors and hyperpriors can be set in the list of priors given to the \code{priors} argument of the \code{disag_model} function.
 In order to print more verbose output the user can set the \code{silent} argument to FALSE.
@@
 %
-<<disag_model, message = FALSE>>=
+<<disag_model, message = FALSE, warnings=FALSE>>=
 fitted_model <- disag_model(data = dis_data,
@@
 The \code{summary} function gives the estimate and standard error of the fixed effect parameters in the model, as well as the negative log likelihood and in-sample performance metrics (root mean squared error, mean absolute error, pearson correlation coefficient and spearman rank correlation coefficient).
+The standard error is calculated using the delta method~\citep{doob1935}, via the \code{sdreport} method in the \pkg{TMB} package.
 The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions, as shown in Figure~\ref{fig:plotmodel}.
@@
 The \code{plot} function can be used on the \code{disag_prediction} objects, as shown in Figure \ref{fig:plotprediction}.
+From Figure~\ref{fig:plotdata}, it can be seen that the polygon response data does not include the islands, whereas the covariate rasters do.
+As can be seen in Figure~\ref{fig:plotprediction}, out-of-sample malaria predictions of these islands have been made, however these predictions do not contribute to the in-sample performance in Figure~\ref{fig:plotmodel}.
 
@@
 Using three simple functions within the \pkg{disaggregation} package we have been able to fit a Bayesian spatial disaggregation model and predict pixel-level incidence rate across Madagascar using aggregated incidence data and pixel-level environmental covariates.
+
+The same technique can be used for spatio-temporal disaggregation modelling.
+In order to achieve this, dynamic covariates are required, for example, annual covariate rasters, as well as a spatio-temporal field, which could be achieved by adding a time varying autoregressive component to the spatial field.
+A difficulty with this approach is that it results in significant increase in computational complexity of the optimisation problem, most notably from the spatio-temporal field, leading to a significant increase in the time taken to perform the optimisation.
+The package can be developed to include spatio-temporal disaggregation models, however that is beyond the scope of this paper.
 
@@
 This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations, doubling the number of iterations each time until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
+The property $\hat{R}$ is known as the Gelman-Rubin diagnostic~\citep{Rhat}, and is commonly used to evaluate MCMC convergence across multiple chains.
 In contrast, fitting the model using the Laplace approximation via \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
 Fitted parameter values for both of these methods are given in Table~\ref{tab:mcmcresults}.
+It can be seen that the model parameters are very similar between the two methods.
+It is worth noting that the field parameters given in the Table are $log(\rho)$ and $log(\sigma)$.
+If we transform these parameters, the values given in Table~\ref{tab:mcmcresults} correspond to a difference in the range of the field from 1.9 (TMB) to 2.3 (MCMC), and a difference in the standard deviation of the field from 1.0 (TMB) and 1.06 (MCMC).
+Considering this along with the large accompanying uncertainties, these field parameters are very similar.
 

### 2021-01-23 article.Rnw
More changes from review responses

@@
 Disaggregation modelling is unorthodox as the predictions are at a different scale to the response data, i.e., the number of rows in the covariate data is different to that of the response data.
-The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to be very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless}.
+The spatial modelling software package, \pkg{INLA}~\citep{rue2009approximate}, or integrated nested Laplace approximation, has been shown to be very useful in a wide variety of circumstances, however it is not flexible enough for the unorthodox nature of the disaggregation problem except in the special case of the linear link function \citep{wilson2017pointless, moraga2017geostatistical}.
 Disaggregation models can be implemented in \pkg{TMB}~\citep{tmb2016}, or template model builder, with a lot of flexibility, however the data manipulation required to format the model objects and construct the model definition in C++ is non-trivial.
@@
 The data, $y_i$, are assumed to be created by the aggregation of the counts over the polygon, i.e., the count data of the polygon is given by the sum of the count data for all the pixels within that polygon.
-As well as the model covariates, an aggregation raster can be provided transform from the pixel level predictions (rate) to count data, from which it is trivial to calculate the corresponding polygon value by summing up the values for the individual pixels within the polygon.
+An aggregation raster must be provided to transform from the pixel level predictions (rate) to count data, from which it is trivial to calculate the corresponding polygon value by summing up the values for the individual pixels within the polygon.
 For example, in epidemiology, we may have as our response the number of people that contract a certain disease in a given period of time (case incidence).
-Our rate would be the incidence rate of cases per population, where the aggregation raster corresponds to population.
-Predictions of incident rate per population are made at pixel level, the population (aggregation) raster is used to convert these values to number of cases, which can then be summed to get predicted polygon level cases.
+Our rate would be the number of cases per population, where the aggregation raster corresponds to population.
+The case generating process is modeled at the pixel level, with these processes then aggregated to obtain the likelihood for the aggregated observed data.
 
@@
 with two hyperparameters: $\rho = \frac{\sqrt{8\nu}}{\kappa}$, the range of the field (beyond which correlation is $<$ 0.1), and $\sigma$, the marginal standard deviation, relating to the magnitude of the variation in the field.
-The parameter $\nu$ is the smoothness and is fixed at 2.
-The Matern field is at least $m$ times mean square differentiable if and only if $\nu > m$~\citep{stein1999interpolation}, therefore $\nu = 2$ is chosen as it is the smallest integer (least smooth) for the field to be differentiable.
+The parameters $\rho$, $\sigma$ and $\nu$ are very difficult to identify together, therefore we fix $\nu$, to be able to identify $\rho$ and $\sigma$.
+The parameter $\nu$ is the smoothness and is fixed at 1, as it is considered the more natural basic choice for two-dimensional ($d=2$) models~\citep{INLA}.
 The Gamma function is given by $\Gamma(\nu) = (\nu - 1)!$ and $K_{\nu}$ is the modified Bessel function of the second kind.
@@
 
-The predictions are made at pixel level, however the response data exists as count data at the polygon level.
-Therefore, to calculate the contribution  to the likelihood from the data we must aggregate the pixel predictions, (via the aggregation raster $a_{ij}$, to get the predicted polygon count.
+Incidence is modeled at the pixel level, and the response data exists as count data at the polygon level.
+Therefore, to calculate the likelihood we must aggregate the pixel rate, via the aggregation raster $a_{ij}$, to get the expected polygon count.
+This aggregation is defined by:
 
@@
 \begin{equation} \label{eq:poisson}
-y_i \sim \textrm{Poisson}(\textrm{cases}_i)
+y_i | \beta_0, \beta_i, \textrm{GP}, u_i \sim \textrm{Poisson}(\textrm{cases}_i)
 \end{equation}
@@
 \begin{equation} \label{eq:gaussian}
-y_i \sim \textrm{Normal}(\textrm{cases}_i, \sigma_i)
+y_i | \beta_0, \beta_i, \textrm{GP}, u_i \sim \textrm{Normal}(\textrm{cases}_i, \sigma_i)
 \end{equation}
@@
 \begin{equation} \label{eq:binomial}
-y_i \sim \textrm{Binomial}(M_i, \textrm{rate}_i)
+y_i | \beta_0, \beta_i, \textrm{GP}, u_i \sim \textrm{Binomial}(M_i, \textrm{rate}_i)
 \end{equation}
@@
 The binomial model could be used when $y_i$ is the prevalence of a disease in a sample of $M_i$ people in the polygon.
+
+The pixel predictions of incident rate are calculated from the fitted model parameters using Equation~\ref{eq:lp}.
 
@@
 This is chosen to be consistent with the prior set by INLA for the dispersion of the normal likelihood.
+We aim for consistency with INLA as INLA is one of the most commonly used packages for Bayesian spatial modeling and due to the parallels in computation approach such as the use of a Laplace approximation and the SPDE approximation to the Gaussian Random Field.
 
@@
 
+TMB is a package which allows the user to define and fit latent variable models.
+The user defines the objective function (typically the likelihood or posterior density function) in C++ and TMB then implements the Laplace approximation to integrate out random effects and generates functions for evaluating the objective and gradient (via automatic differentiation).
+These functions can then be passed to an optimiser (for maximum likelihood estimation or to find the posterior maximum) or sampler (for MCMC) in R.
+\pkg{TMB} can fit both Frequentist and Bayesian models.
+In this case we have chosen the Bayesian approach as it is most widely used in the spatial modelling community.
+
 The \pkg{disaggregation} package contains a C++ function that defines the model and computes the joint likelihood as a function of the parameters and the random effects, in the format expected by \pkg{TMB}.
@@
 Evaluation of the objective function and its derivatives is performed via \proglang{R}.
-
-\pkg{TMB} can fit both Frequentist and Bayesian models. In this case we have chosen the Bayesian approach as it is most widely used in the spatial modelling community.
 
@@
 
+In this paper we compare our method to Markov chain Monte Carlo (MCMC).
+MCMC methods involve constructing Markov chains which (after a sufficient warm up period) produce samples from the desired probability distribution.
+In high dimensional problems, these methods can be extremely slow as many steps are needed to converge to and effectively sample from the target distribution.
+This is particularly problematic when computing the likelihood is costly, as the likelihood is typically evaluated at each step.
+In spatial settings with a Gaussian field, MCMC becomes infeasible for disaggregation modelling.
+Our method presented in the paper leverages the Laplace approximation to provide a much faster way of fitting disaggregation models.
+
 \section{Package usage} \label{sec:code}
@@
 Therefore, such environmental covariates would be useful in a malaria disaggregation model to inform fine-scale distributions.
-In this model we use the environmental covariates of mean land surface temperature (LST), elevation, and enhanced vegetation index (EVI) to inform spatial heterogeneity in malaria risk.
+In this model we use the environmental covariates of mean land surface temperature (LST), elevation, and enhanced vegetation index (EVI), at a resolution of approximately $5\times5~km$, to inform spatial heterogeneity in malaria risk.
 These covariates are obtained from the Moderate Resolution Imaging Spectroradiometer (MODIS), which provides many measurements over the entire Earth's surface (\url{https://modis.gsfc.nasa.gov/data/}).
@@
 In this case our pixel predictions correspond to malaria incidence rate, so we use population to aggregate pixel incidence rate by summing the number of cases (rate weighted by population).
-Raster surfaces of population for the years 2010 and 2015, were created using data from WorldPop~\citep{tatem2017worldpop} and from GPWv4~\citep{GpwNasa} where WorldPop did not have values.
+Raster surfaces of population for the years 2010 and 2015 at a resolution of approximately $5\times5~km$, were created using data from WorldPop~\citep{tatem2017worldpop} and from GPWv4~\citep{GpwNasa} where WorldPop did not have values.
 The population raster for 2016 was created by linear interpolation between 2010 and 2015 and extending out to 2016.
@@
 Covariate rasters or population rasters may be accompanied by significant uncertainty.
-There is no established method within the \pkg{disaggregation} package to deal with this.
-In order to propagate this uncertainty to the predictions, users may run the model many times, varying these rasters within the known uncertainty.
-Both the range in predictions for different models and the model uncertainty within a model allows a quantification of the overall uncertainty of predictions.
+There is no robust way within the package to propagate this uncertainty.
+However, if the user were able to sample from the covariate rasters and run many instances of the model, they could estimate an overall uncertainty.
+Whether this is a good approximation of the true uncertainty will depend on the specific use case.
 
 The covariate rasters and aggregation raster provided must be of the same spatial scale and must be spatially aligned.
-For this example the rasters used are at a resolution of approximately $5\times5~km$ across Madagascar.
 For spatially misaligned rasters, there are packages in \prolang{R} to align rasters such as \code{align_rasters} in the \pkg{gdalUtils} package~\cite{gdalutils}.
@@
 The function \code{getCovariateRasters} is a helper function contained within the \pkg{disaggregation} package.
-The user provides a directory name and shapefile, it finds all the raster files within that directory and creates a RasterStack with the same extend as the shapefile provided.
+From a user defined a directory name and shapefile, all the raster files within that directory are read and a RasterStack is created with the same extent as the shapefile provided.
 In this example we are building a RasterStack of the covariate rasters with the same extent as the population raster.
@@
 The aggregation raster defines how the pixels within each polygon are aggregated.
-The disaggregation model performs a weighted sum of the pixel predictions, weighted by the pixel values in the aggregation raster.
+The disaggregation model performs a weighted sum of the pixel predictions, weighted by the pixel values in the aggregation raster, as shown in~Equation~\ref{eq:agreggate1}.
 In this case our pixel predictions are malaria incidence rate, so we use the population raster to aggregate pixel incidence rate by summing the number of cases (rate weighted by population).
@@
 
-To fit a model with a Gaussian field we approximate a Gaussian field using Gaussian Random Markov Fields (GRMF), and solve using the stochastic partial differential equation (SPDE) approach~\citep{Lindgren2011}.
-This reduces the complexity of the problem to a sparse precision matrix, allowing explicit computation of the GMRF representation.
+To fit a model with a Gaussian field we approximate a Gaussian field using Gaussian Markov Random Fields (GMRF), and solve using the stochastic partial differential equation (SPDE) approach~\citep{Lindgren2011}.
 This approach requires building a mesh, i.e.
 splitting the space into finite elements, and calculating field values at the mesh nodes.
-If a mesh is too fine, the field values take too long to compute, however if the mesh if too coarse, the resulting field is a poor approximation and leads to mesh artifacts in the predictions.
+If a mesh is too fine, the field values take too long to compute, however if the mesh is too coarse, the resulting field is a poor approximation and leads to mesh artifacts in the predictions.
 The 2D mesh for the spatial field is built using the \code{build_mesh} function, which makes use of the \pkg{INLA} function \code{inla.mesh.2d}.
 The user can control the parameters of the mesh, including the granularity, using the argument \code{mesh.args} in the \code{build_mesh} function.
-The parameters \code{max.edge}, \code{cut} and \code{offset} can be set in the \code{mesh.args} list and are defined within the \code{inla.mesh.2d} function.
-By providing a two values to the \code{max.edge} parameter, the mesh contains an inner region of finer mesh, and an outer coarse region.
-This approach allows the region of interest to have a fine a mesh as necessary without also requiring time consuming computations in the sea regions.
+The parameters \code{max.edge}, \code{cut} and \code{offset} that can be set in the \code{mesh.args} list are defined within the \code{inla.mesh.2d} function.
+By providing two values to the \code{max.edge} parameter, the mesh contains an inner region of finer mesh, and an outer coarse region.
+This approach allows the region of interest to have a fine as mesh as necessary without also requiring time consuming computations in the sea regions.
 The mesh can take several minutes to construct, so to prepare the data without building the mesh the user can set the \code{makeMesh} flag to FALSE.
@@
 \item \code{data}: Object of class \code{disag_data} returned by the \code{prepare_data} function.
-\item \code{priors}: List of priors to use for the model. For any not set, the default priors will be used.
+\item \code{priors}: List of priors and hyperpriors to use for the model. For any not set, the default priors will be used.
 \item \code{iterations}: Maximum number of iterations the model can run for to find an optimal point.
@@
 The \code{summary} function gives the estimate and standard error of the fixed effect parameters in the model, as well as the negative log likelihood and in-sample performance metrics (root mean squared error, mean absolute error, pearson correlation coefficient and spearman rank correlation coefficient).
-The standard error is calculated using the delta method~\citep{doob1935}, via the \code{sdreport} method in the \pkg{TMB} package.
+The standard error is calculated based on the approximate posterior from the \pkg{TMB} package.
 The \code{plot} function produces two plots: one of the fixed effects parameters and one of the observed data against in-sample predictions, as shown in Figure~\ref{fig:plotmodel}.
@@
 From Figure~\ref{fig:plotdata}, it can be seen that the polygon response data does not include the islands, whereas the covariate rasters do.
-As can be seen in Figure~\ref{fig:plotprediction}, out-of-sample malaria predictions of these islands have been made, however these predictions do not contribute to the in-sample performance in Figure~\ref{fig:plotmodel}.
+As can be seen in Figure~\ref{fig:plotprediction}, out-of-sample malaria predictions of these islands have been made, however these predictions do not contribute to the in-sample performance in Figure~\ref{fig:plotmodel}, nor to cross validation performance.
+In fact, these islands are not part of Madagascar at all, they are the islands of Comoros and Mayotte.
 
@@
 It is important to note that this is a useful feature of the \pkg{disaggregation} package, to be able to create the \pkg{TMB} model object (using the \code{disag_model} function with one iteration) and pass it directly to \pkg{tmbstan}.
-The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 71 hours.
+The model is fitted by running the MCMC algorithm for 8000 iterations with 2000 of those as warmup, which took 74 hours.
 This number of iterations was chosen by running the MCMC algorithm repeatedly, starting at 1000 iterations, doubling the number of iterations each time until the value of the MCMC convergence statistic, $\hat{R}$, dropped below 1.05 for all model parameters.
@@
 In contrast, fitting the model using the Laplace approximation via \pkg{TMB} within the \pkg{disaggregation} package took 56 seconds.
+
 Fitted parameter values for both of these methods are given in Table~\ref{tab:mcmcresults}.
@@
 It is worth noting that the field parameters given in the Table are $log(\rho)$ and $log(\sigma)$.
-If we transform these parameters, the values given in Table~\ref{tab:mcmcresults} correspond to a difference in the range of the field from 1.9 (TMB) to 2.3 (MCMC), and a difference in the standard deviation of the field from 1.0 (TMB) and 1.06 (MCMC).
-Considering this along with the large accompanying uncertainties, these field parameters are very similar.
+If we transform these parameters to their natural form, the values given in Table~\ref{tab:mcmcresults} correspond to a difference in the range of the field, $\rho$, from $1.9 \pm 0.6$ (TMB) to $2.3 \pm 0.8$ (MCMC), and a difference in the standard deviation of the field, $\sigma$, from $1.00 \pm 0.17$ (TMB) and $1.07 \pm 0.21$ (MCMC).
+Considering these values, along with their large accompanying uncertainties, these field predictions are very similar between the two methods.
+In general, the Laplace approximation will perform worse when the posterior is less Gaussian; this occurs when the prior has more influence in the model, for example, when there is less data or for hierarchical parameters.
 
@@
 \begin{tabular}{l|ll|ll}
- & \multicolumn{2}{|c|}{MCMC (71 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
+ & \multicolumn{2}{|c|}{MCMC (74 hours)} & \multicolumn{2}{|c}{TMB (56 seconds)} \\
 Parameter & Mean & SD & Mean & SD \\
 \hline
-Intercept & -3.22 & 0.33 & -3.20 & 0.27 \\
+Intercept & -3.23 & 0.31 & -3.20 & 0.27 \\
 Slope 1 & -0.47 & 0.20 & -0.47 & 0.20 \\
-Slope 2 & 0.41 & 0.22 & 0.39 & 0.22 \\
-Slope 3 & 0.21 & 0.29 & 0.21 & 0.29 \\
-$log(\tau_u)$ & 0.90 & 0.27 & 0.98 & 0.27 \\
-$log(\sigma)$ & 0.062 & 0.009 & -0.002 & 0.166 \\
+Slope 2 & 0.41 & 0.23 & 0.39 & 0.22 \\
+Slope 3 & 0.21 & 0.31 & 0.21 & 0.29 \\
+$log(\tau_u)$ & 0.92 & 0.28 & 0.98 & 0.27 \\
+$log(\sigma)$ & 0.07 & 0.20 & -0.002 & 0.166 \\
 $log(\rho)$ & 0.82 & 0.35 & 0.63 & 0.33 \\
@@
 The \pkg{disaggregation} package produces similar results to the MCMC algorithm.
-However the model fitting using the \pkg{disaggregation} package took 56 seconds in contrast to the 71 hours taken for the MCMC run.
+However the model fitting using the \pkg{disaggregation} package took 56 seconds in contrast to the 74 hours taken for the MCMC run.
 Therefore, it can be seen that the \pkg{disaggregation} package provides a quick and simple way to run disaggregation models, that can be prohibitively slow using MCMC.
@@
 \includegraphics[width=\linewidth]{figures/mcmc_trace.pdf}
-\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 71 hours. It was run for 8000 iterations with 2000 of those as warmup.}
+\caption{Trace of the fixed effects parameters for MCMC using NUTS sampling, running the algorithm for 74 hours. It was run for 8000 iterations with 2000 of those as warmup.}
 \label{fig:mcmclong}
@@
 
+\section*{Data availability}
+
+All data used in this paper, including the results of the MCMC algorithm, can be found here: \url{https://figshare.com/projects/disaggregation/96095}.
+
 \section*{Acknowledgments}
